### 2024-01-31 pnrs.tex
lipics compiles

%! TeX program = xelatex
%! TeX root = paper-name.tex
%! lang = en-US
\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}

% INPUTS
\input{globals/packages}
\input{globals/colorscheme}

% CONFIG
\input{globals/macros}
\input{globals/mathdefs}
\input{globals/knowledges.kl}

\usepackage{comment}
%\includecomment{forshort}
%\excludecomment{proof}

\crefname{fact}{Fact}{Facts}
\Crefname{fact}{Fact}{Facts}
\crefname{algocf}{Algorithm}{Algorithms}
\Crefname{algocf}{Algorithm}{Algorithms}

% DOCUMENT
\begin{document}
\input{globals/metadata}

% FRONT MATTER
\maketitle
%\makeabstract
%\acknowledge

\input{parts/intro.tex}
\input{parts/prelim.tex}
\input{parts/counter.tex}
\input{parts/correction.tex}
\input{parts/starfree.tex}
\input{parts/transducers.tex}
\input{parts/conclusion.tex}

% BACKMATTER
\bibliographystyle{plainurl}
\bibliography{globals/papers}

\appendix

\include{parts/appendix.tex}

\end{document}