### 2024-02-12 parts/transducers.tex
big big rewrite

%! TeX program = xelatex
%! TeX root = ../pnrs.tex
%! lang = en-US
\section{Beyond commutativity}
\label{beyond-commutative:sec}

\AP The goal of this section is to provide tools that do not require the
assumption of \kl{commutativity}. To that end, we introduce the notion of
\kl{$k$-residual transducer}, that generalizes the notion of residual
transducer introduced for \kl{$\Rel$-polyregular functions} in \cite{LOPEZ23b}.
This notion of residual transducer has its own interest, as it shifts the
attention from (compatible) equivalence relations of finite index that are the
classical tool in automata theory, towards (compatible) order relations that
are \kl{well-quasi-ordered}, which is the order counterpart of having finite
index. Let us recall that a sequence $\seqof{u_i}{i \in \Nat}$ of elements in a
quasi-ordered set $(X, \leq)$ is \intro{good} whenever there exist $i < j$ such
that $u_i \leq u_j$. The set $X$ is a \intro{well-quasi-ordering} when every
infinite sequence is \reintro{good}. A sequence is \intro{bad} when it is not
\reintro{good}. The notion of \reintro{good sequences} can also be applied to
binary relations that are not orderings, and a binary relation $R$ for which
every infinite sequence is \reintro{good} is said to be \intro{well}
\cite{MELL98}.

The following definition of transducers is based on
the marble transducers \cite{EHB99}, and follows the general pattern of
transducers introduced in \cite{LOPEZ23b}.

\begin{definition}
    Let $\mathcal{H}$ be a family of functions
    from $\Sigma^*$ to $\Nat$.
    An \intro{$\mathcal{H}$-transducer} $\aTransd$ is
    a tuple $(Q, q_0, \delta, \lambda, F)$ where
    \begin{itemize}
        \item $Q$ is a finite set of states,
        \item $q_0 \in Q$ is an initial state,
        \item $\delta \colon Q \times \Sigma \to Q$
            is a deterministic transition function,
        \item $\lambda \colon Q \times \Sigma \to \mathcal{H}$
            is a correction function,
        \item $F \colon Q \to \Nat$ is an output function.
    \end{itemize}
\end{definition}

Given the syntax of an \kl{$\mathcal{H}$-transducer}, let us now 
introduce its semantics, which is given by a simple recursion
scheme.
\begin{definition}
    Let $\aTransd \defined (Q,q_0, \delta, \lambda, F)$ be an \kl{$\mathcal{H}$-transducer}.
    The function
    $\aTransd \colon Q \times \Sigma^* \to \Nat$
    is defined inductively  as follows:
    \begin{itemize}
        \item $\aTransd(q, \varepsilon) \defined F(q)$;
        \item $\aTransd(q, a u) \defined \aTransd(\delta(q,a), u)
            + \lambda(q,a)(u)$.
    \end{itemize}
    The function \intro{computed} by a transducer $\aTransd$
    is $w \mapsto \aTransd(q_0, w)$.
\end{definition}

\begin{example}
    \label{non-canonical-transd:ex}
    Let $f \colon \set{a}^* \to \Nat$
    be defined as
    $f(\varepsilon) = 1$,
    $f(a) = 0$,
    and $f(a^n) = n - 1$ for all $n \geq 2$.
    \textbf{TODO: split the example}
    $f(\varepsilon) = 1$,
    $f(a) = 0$,
    $f(a^2) = 1$,
    and $f(a^n) = n - 3$ for all $n \geq 3$.
    The function is \kl{computed} by the two automata
    in \cref{non-canonical-transd:fig}.
\end{example}

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.49\linewidth}
        \begin{tikzpicture}[
            etat/.style={minimum size=2em}
            ]
            \node[etat,state,initial,
                accepting by arrow,
                accepting text={$1$},
                accepting where=below,
                ] (A) at (0,0) {$\varepsilon$};
            \node[etat,state,
                accepting by arrow,
                accepting text={$0$},
                accepting where=below,
                ] (B) at (2,0) {$a$};

            \draw[->] (A) to node[midway, below] {$a \mid 0$} (B);

            \draw[->] (B) to[bend right=45] node[midway, above] {$a \mid 1 + \ind{\card{w} \geq 1}$} (A);
        \end{tikzpicture}
        \caption{This automaton has a \kl{counter}.}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\linewidth}
        \begin{tikzpicture}[
            etat/.style={minimum size=2em}
            ]
            \node[etat,state,initial,
                accepting by arrow,
                accepting text={$1$},
                accepting where=below,
                ] (A) at (0,0) {$\varepsilon$};
            \node[etat,state,
                accepting by arrow,
                accepting text={$0$},
                accepting where=below,
                ] (B) at (2,0) {$a$};

            \draw[->] (A) to node[midway, below] {$a \mid 0$} (B);

            \draw[->] (B) edge[loop right] node[midway, right]
                {$a \mid 1 + \ind{\card{w} = 0}$} (B);
        \end{tikzpicture}
        \caption{This automaton is \kl{counter-free}.}
    \end{subfigure}
    \caption{Two \kl{$\NPoly[0]$-transducers}
    \kl{computing} the function of \cref{non-canonical-transd:ex}.}
    \label{non-canonical-transd:fig}
\end{figure}

\AP As in the case of finite automata, we generalize the function $\delta
\colon Q \times \Sigma \to Q$ to words by defining $\delta(q, \varepsilon) =
q$, and $\delta(q,au) = \delta(\delta(q,a), u)$. Similarly, we write
$\lambda(q,\varepsilon) \defined 0$, and recursively define $\lambda(q, au)
\defined \app{\lambda(q,a)}{u} + \lambda(\delta(q,a), u)$. Following these
equations, we can write for all $q \in Q$, $u,v \in \Sigma^*$, $\aTransd(q, uv)
= \aTransd(\delta(q, u), v) + \lambda(q, u)(v)$. Finally, let us notice that if
$\mathcal{H}$ is closed under finite sums and shifts, then $\lambda(q, u) \in
\mathcal{H}$ for all $(q,u) \in Q \times \Sigma^*$.

\AP Let us recall that a \intro{counter} in an automaton is a pair $u,w$ of
words together with an integer $n > 1$, such that $\delta(q_0, uw^n) =
\delta(q_0, u)$ and $\delta(q_0, u w^k) \neq \delta(q_0, u)$ for all $k < n$.
An automaton is \reintro{counter-free} if it contains no \reintro{counters}.
This notion can be applied to \kl{$\mathcal{H}$-transducers} by considering
counters of their underlying automata. It is already known in the literature
that this notion of transducers correctly captures both $\NSF$ and $\NPoly$ in
the sense that the following fact will clarify.

\begin{lemma}[{\cite{gaetanphd}}]
    \label{transducer-nsf-npoly:lemma}
    Let $f \colon \Sigma^* \to \Nat$, and let $k \in \Nat$.
    The following are equivalent:
    \begin{itemize}
        \item $f \in \NPoly[k]$ (respectively $\NSF[k]$),
        \item $f$ is \kl{computed} by an \kl{$\NPoly[k-1]$-transducer}
            (respectively by an \kl{$\NSF[k-1]$-transducer} that is 
            \kl{counter-free}).
    \end{itemize}
\end{lemma}

\AP The core concept of this section is that of a \intro{residual} of a
function $f \colon \Sigma^* \to \Rel$,  defined by $\app{f}{u} \defined w
\mapsto f(uw)$. The collection of \reintro{residuals} of a function $f$ is
denoted $\Res(f)$ and is defined as the set of $\app{f}{u}$ where $u$ ranges
over words in $\Sigma^*$. Let $k \in \Nat$. We define $f \zpolyequiv[k] g$ if
and only if $(g - f) \in \ZPoly[k-1]$, and $f \npolyleq[k] g$ if and only if
$(g - f) \in \NPoly[k-1]$. Given a function $f \colon \Sigma^* \to \Rel$, we
are going to study both $(\Res(f), \zpolyequiv[k])$, and $(\Res(f),
\npolyleq[k])$. However, to simplify notations, instead of writing $\app{f}{u}
\npolyleq[k] \app{f}{v}$ when $u,v \in \Sigma^*$, we will use the convenient
notation $u \resleq{f}{k} v$, and directly consider the space $(\Sigma^*,
\resleq{f}{k})$.

\begin{fact}
    \label{good-residual-ordering:fact}
    Let $k \in \Nat$, and let $f \colon \Sigma^* \to \Nat$. Then,
    $(\resleq{f}{k})$ is a quasi-ordering, satisfying the following
    extra properties:
    \begin{enumerate}
        \item For all $u,v,w \in \Sigma^*$, $u \resleq{f}{k} v$
            implies $uw \resleq{f}{k} vw$,
        \item If $u \resleq{f}{k} v$ and $v \resleq{f}{k} u$,
            then $\app{f}{u} = \app{f}{v}$.
    \end{enumerate}
\end{fact}

Given a function $f$, our goal is to leverage $\resleq{f}{k}$ to build a
canonical \kl{$\NPoly[k-1]$-transducer} that \kl{computes} $f$. The naïve
approach is to consider as states some representatives for the $\resleq{f}{k}$
minimal elements of $\Sigma^*$ (assuming there are finitely many of them), all
computing the same function (\cref{good-residual-ordering:fact}), and define
transitions by letting $\delta(u, a)$ be \emph{one} state $v$ such that $v
\resleq{f}{k} ua$. This naïve approach does not yield a canonical model which
is corrected by the following definition.

\begin{definition}
    \label{residual-transducer:def}
    Let $f \colon \Sigma^* \to \Nat$ and $k \in \Nat$.
    A transducer $\aTransd \defined (Q, q_0, \delta, \lambda, F)$
    is a \intro{$k$-residual transducer}
    of $f$ 
    when
    it is a \kl{$\NPoly[k-1]$-transducer}
    satisfying the following properties:
    \begin{enumerate}
        \item $\aTransd$ \kl{computes} $f$;
        \item $Q \subseteq \Sigma^*$ is a \kl{downwards closed}
            for $\prefleq$;
        \item $q_0 = \varepsilon$;
        \item every state $q \in Q$ is accessible from $q_0$;
        \item For all $u, a \in Q$,
            $\delta(u,a)$ is the $\prefleq$-maximal $v \in Q$
            such that $v \prefleq ua$, and $v \resleq{f}{k} ua$.
        \item For all $u,a \in Q$,
            $\lambda(u,a) = \app{f}{ua} - \app{f}{\delta(u,a)}$.
    \end{enumerate}
\end{definition}

Because of the restriction on prefixes, the \kl{$k$-residual transducer}
is actually unique when it exists.
We now introduce \cref{residual:algo} to compute a \kl{$k$-residual transducer}
given a function $f$. Notice that this algorithm requires the ability to test
if a function belongs to $\NPoly[k]$, which is only known to be feasible for
\kl{commutative} polyregular functions. However, the termination of this
algorithm also proves the existence of the \kl{$k$-residual transducer}.

\begin{algorithm}
    $Q \defined \{ \varepsilon \}$;
    $O \defined \setof{ a }{ a \in \Sigma}$;
    $\delta \defined \emptyset$;
    $\lambda \defined \emptyset$;
    $F \defined \emptyset$;

    \While{$O \neq \emptyset$}{
        choose $ua \in O$;

        $O \defined O \setminus \set{ ua }$;

        \eIf{$\exists v \in Q, v = \max_{\prefleq} \setof{w \in Q}{w \prefleq ua \wedge w \resleq{f}{k} ua}$}{
            $\delta(u, a) \defined v$;

            $\lambda(u, a) \defined \app{f}{ua} - \app{f}{v}$;
        }{
            $Q \defined Q \uplus \set{ ua }$;

            $\delta(u,a) \defined ua$;

            $\lambda(u,a) \defined 0$;

            $O \defined O \cup \setof{ uab }{b \in \Sigma}$;
        }
    }
    \For{$u \in Q$}{
        $F(u) \defined f(u)$;
    }
    \Return{$(Q, \varepsilon, \delta, \lambda, F)$};
    \caption{Computing a $k$-residual transducer given a function $f$.}
    \label{residual:algo}
\end{algorithm}

\begin{lemma}
    \label{correct-residual:lemma}
    If \cref{residual:algo} terminates on 
    an input $f \colon \Sigma^* \to \Nat$, $k \in \Nat$,
    then it computes the \kl{$k$-residual transducer} of $f$.
\end{lemma}

\begin{lemma}
    \label{wqo-implies-termination:lemma}
    Let $f \colon \Sigma^* \to \Nat$, and $k \in \Nat$ be such that
    every infinite, $\prefleq$-increasing sequence is \kl{good}
    in $(\Sigma^*, \resleq{f}{k})$
    (or equivalently, such that the relation $(\prefleq \Rightarrow \resleq{f}{k})$
    is \kl{well}).
    Then, \cref{residual:algo} terminates on the input $(f,k)$.
\end{lemma}

The following \cref{n-poly-k-implies-wqo:lemma} is the core argument of the
section, and states that for a function $f \in \NPoly[k]$, not only $(\prefleq
\Rightarrow \resleq{f}{k})$ is \kl{well}, but the quas-ordering $(\resleq{f}{k})$ is a
\kl{well-quasi-ordering}. This has to be compared with the property of being
\kl{$(k,\Nat)$-combinatorial}, which states that ultimately, the function is
increasing in every direction.

\begin{lemma}
    \label{n-poly-k-implies-wqo:lemma}
    Let $k \in \Nat$, and let $f \in \NPoly[k]$.
    Then, $(\Sigma^* \resleq{f}{k})$ is a \kl{well-quasi-ordering}.
\end{lemma}

\begin{theorem}
    \label{non-commutative-npoly:thm}
    Let $f \in \ZPoly$ be a non-negative function, 
    and $k \in \Nat$,
    the following are equivalent:
    \begin{enumerate}
        \item \label{n-poly-1-transd:item} $f$ is \kl{computed}
            by a \kl{$\NPoly[k-1]$-transducer};
        \item \label{n-poly-k:item} $f \in \NPoly[k]$;
        \item \label{n-poly-wqo:item} $(\Sigma^*, \resleq{f}{k})$ is a
            \kl{well-quasi-ordering};
        \item \label{n-poly-well:item} every $\prefleq$-increasing sequence
            of $\Sigma^*$  is a \kl{good sequence}
            for $\resleq{f}{k}$;
        \item \label{n-poly-residual:item} $f$ is \kl{computed} by a
            \kl{$k$-residual transducer}.
    \end{enumerate}
    If $f$ is assumed to be \kl{commutative}, then all the above
    properties are decidable, and the conversions effective.
\end{theorem}
\begin{proof}
    \cref{n-poly-1-transd:item} implies \cref{n-poly-k:item} by
    \cref{transducer-nsf-npoly:lemma}. Then,
    \cref{n-poly-k:item} implies \cref{n-poly-wqo:item} by
    \cref{n-poly-k-implies-wqo:lemma}.
    The implication \cref{n-poly-wqo:item} $\Rightarrow$ \cref{n-poly-well:item}
    is obvious.
    Then, \cref{wqo-implies-termination:lemma} proves
    that \cref{n-poly-well:item} implies \cref{n-poly-residual:item}.
    Finally, because a \kl{$k$-residual transducer} is a \kl{$\NPoly[k-1]$-transducer},
    \cref{n-poly-residual:item} implies \cref{n-poly-1-transd:item}.
\end{proof}