### 2024-02-14 globals/metadata.tex
update ICALP

\date{\today}

\title{Commutative $\Nat$-polyregular functions}

\author{Aliaume Lopez}%
       {University of Warsaw}%
       {ad.lopez@uw.edu.pl}%
       {https://orcid.org/0000-0002-1825-0097}%
       {}

\authorrunning{A. Lopez}

%\titlerunning{Commutative $\Nat$-polyregular functions}
\date{\today}

\Copyright{Aliaume Lopez}

%TODO mandatory; please add comma-separated list of keywords
\keywords{Rational series, weighted automata, polyregular function, commutative input} 
\category{} %optional, e.g. invited paper
\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website

\ccsdesc[500]{Theory of Computation~Models of Computation}
\ccsdesc[500]{Theory of computation~Automata extensions}

\newcommand{\acknowledge}{
    Acknowledge Gaëtan
}

\newcommand{\makeabstract}{
    \begin{abstract}
This paper addresses two questions regarding $\Nat$-polyregular functions, that
forms a proper subset of $\Nat$-rational series. We show that given a $\Rel$-rational
series, it is decidable whether it is computable via a commutative
$\Nat$-polyregular function, and provide a counter-example to the theorem of
Karhumäki that studied the same question in the case of polynomials. We also
prove that it is decidable whether a commutative $\Nat$-polyregular function is
star-free, by proving the stronger statement that star-free $\Rel$-polyregular
functions that are $\Nat$-polyregular are in fact computable using a star-free
$\Nat$-polyregular function. Building towards answering the same questions in the
non-commutative case, we present a canonical model of computation of
$\Nat$-polyregular functions by generalizing the notion of residual transducers
previously introduced in $\Rel$-polyregular functions.
    \end{abstract}
}

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{ICALP 2023}
\EventAcronym{ICALP}
\EventYear{2023}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{285}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%