### 2024-03-27 reviews/rebuttal.md
review

---
title: Rebuttal
author: Aliaume LOPEZ
date: 2024-03-27
lang: en-GB
---

We would like to thank the reviewers for their comments and suggestions, which
are all constructive and will definitely help to improve the quality of the
paper. In particular, we will add the reference to the ArXiv paper
`arXiv:2301.09234` in the introduction which is definitely relevant.

Before answering to the questions of the reviewers, we would like to point out
that the summary of Reviewer 1 is slightly misleading as the paper does not
manage to decide whether a function is "$\mathbb{N}$-polyregular" in general,
but only does so in the case where the function is assumed to be *commutative*
(which is a strong restriction). Without the assumption of commutativity, the
problem remains open. We will make sure to put more emphasis on this point.

---

It is our understanding that a significant part of the reviewers' concerns
relates to the high number of typos and grammatical errors in the manuscript.
We sincerely apologize for these mistakes that made reading this paper
difficult, and we acknowledge the lack of consistency in the bibliographic
section and in particular the mix of unicode and non unicode bibtex entries
which butchered the name of Lê Thành Dũng Tito Nguyên. We will definitely
correct these issues in the revised version of the manuscript. We decided to
not respond to specific typos and grammatical errors in this rebuttal and focus
on more global issues.

---

A second point raised by the reviewers is the lack of clarity for non-expert
readers, with a technical introduction which does not take enough time to
explain the state of the art, and for instance clearly state results obtained
in [9] to compare them with what is done in this paper. To address this issue,
we will take more time to introduce historical results and explain the various
classes of functions that have been studied before, expanding Figure 1 with
concrete examples. We will take time to introduce the various notions of
string-to-string functions (mealy machines, sequential transducers, rational
functions, regular functions, polyregular functions) with a clear table
comparing them and stating the status of "deciding aperiodicity" for each of
these classes.

However, it seems reasonable for us to assume that the reader is familiar at
least with finite monoids, monadic second order logic on words, and weighted
automata (hence, with rational series) which are well established theories:
describing them in a few sentences with a reference to a standard book should
already greatly increase the readability of the paper.

The informal uses of notions such as "growth rate" and "commutativity" in the
introduction were intended to provide intuitions to the reader, but we
understand that it was counterproductive and will be more careful in the
revised version. This is not an isolated issue, and we will make sure to avoid
such informal statements.

We also understand that whether a proof belongs to the appendix or the main
body of the paper felt arbitrary, and that the key ingredients of the proofs
where not clearly identified. While it is difficult to answer to such a comment
without rewriting the paper, we will try to illustrate on a few examples.

- Lemma 7 is a simple proof in the case of polyregular functions, and has
  a sligthly more technical proof[^1] in the case of rational series, but the
  main interest of the paper is in polyregular functions, and therefore the
  proof is in appendix.
- Lemma 19 is a technical result which while technically new, follows proof
  schemes similar to those of [9, 12], and provides another proof of the
  correct part of the theorem of [23]. This is why it is in the appendix.
- Lemma 26 is the actual core of Theorem 29, which is a corrected version of
  a theorem from [23]. In the (incorrect) proof of [23], the error lies in the
  inductive step of the analogue of Lemma 26, which is why we decided to put
  and this is why this particular proof was kept in the main body of the paper,
  while the other technical details about multivariate polynomials were moved
  to the appendix. 
- Lemma 31 has a proof that is a bit technical, but the important part of the
  lemma is actually the statement itself, which decomposes a commutative
  polyregular function into actual *polynomials*. Hence, the proof is in
  appendix.

However, we agree that all the proofs of sections 5 and 6 are in appendix
except for the ones that combine technical lemmas, and a better trade-off
between the sections will be found in the revised version of the paper. We
believe that the interesting proofs are those of Lemmas 44, and 45, which
really leverage the notion of well-quasi-ordering.

We will therefore revise the "Outlook" section to shape it more as a discussion
of the results and their implications, which will leave space for a more
detailed introduction, explanations of the results, and interesting proofs of
section 5.

---

To conclude on the relationship between this paper and [9], which should be
made clearer, let us try to replace both papers in context.

The concern of [9] is about $\mathbb{Z}$-polyregular functions (which are
a subclass of $\mathbb{Z}$-rational series), defining a notion of
"aperiodicity" for such functions, and proving that aperiodicity is decidable.
The work of [9] could be put into context as follows: since the introduction of
polyregular functions [4,5], there has been a corresponding notion of
"aperiodic" polyregular functions. However, decidability of aperiodicity
remains an open question, even in the specific case of unary output (the
function output words over a single letter alphabet). Notice that polyregular
functions with unary output are in fact functions from $\Sigma^*$ to
$\mathbb{N}$ (we write a number in unary).

While the work of [9] could be interpreted as studying $\mathbb{Z}$-rational
series, it is really about deciding aperiodicity of a "variation" of
polyregular functions with unary output that allows "negative outputs".
Unfortunately this "variation" is substantial, and the relationship between
$\mathbb{Z}$-polyregular functions (of [9]) and $\mathbb{N}$-polyregular
functions (unary output) remained unclear up to this paper. In particular the
decidability of aperiodicity for $\mathbb{N}$-polyregular functions could not
be deduced directly from [9].

This paper provides the combinatorial tools to understand the differences
between $\mathbb{Z}$-polyregular functions and $\mathbb{N}$-polyregular
functions (for instance, by giving decision procedures), and leverages these
tools to prove that aperiodicity is decidable for the latter **in the
commutative case**. This is a significant step towards the decidability of
aperiodicity for polyregular functions. The next step would be to drop the
assumption of commutativity, and then to allow the outputs to non-unary
alphabets. Both steps are beyond the scope of the current paper.

We hope that the above clarifications and proposed changes will address the
concerns of the reviewers, and would like to thank them again for the
constructive feedback.

[^1]: There is actually a simpler proof of Lemma 7 in the case of rational
series that avoids the use of Zariski closures and is more direct.