### 2024-03-29 parts/intro.tex
appendix + proofrefs

%! TeX program = xelatex
%! lang = en-US
% MAIN MATTER
\section{Introduction}
\label{introduction:sec}

\AP Traditional results from automata theory state the equivalence of languages
defined in terms of deterministic finite automata and regular expressions
\cite{KLEE56}, Monadic Second Order Logic ($\intro*\MSO$) \cite{BUCH60}, or
finite monoids \cite{SCHU61}. Among regular languages, there exists a robust
subclass sharing similar characterizations, namely the \emph{star-free} regular
languages. These are obtained via counter-free (minimal) automata
\cite{MNPA71}, represented by star-free regular expressions \cite{SCHU65},
definable in First Order Logic ($\intro*\FO$) \cite{PEPI86}, and recognized by
aperiodic monoids \cite{SCHU65}. It is decidable whether a regular language is
aperiodic, thanks to the characterizations in terms of minimal automaton and/or
syntactic monoid.

\AP There have been recurrent attempts to generalize the notions of regularity
from languages to functions, leading to classes of various expressiveness, such
as\footnote{This list is far from exhaustive.} Mealy Machines ($\intro*\Mealy$)
\cite{MEAL55}, sequential functions ($\intro*\Sequential$) \cite{SCHU77},
rational functions ($\intro*\Rational$) \cite{EILE74}, regular functions
($\intro*\Regular$) \cite{ENMA02}, and polyregular functions ($\intro*\Poly$)
\cite{BOKL19}. For all of these models, an \emph{aperiodic} counterpart can be
defined (respectively, $\intro*\AMealy$, $\intro*\ASequential$,
$\intro*\ARational$, $\intro*\ARegular$, $\intro*\SF$), lifting the
correspondence between counter-free automata, star-free languages and aperiodic
monoids to the functional setting
\cite{FKT14,BOJA14,CADA15,DJR16,BDK18,BOKL19,DGK21}. The inclusions between
these classes of functions are all known to be strict, and we depicted in
\cref{computational-models:fig} the status of the \emph{membership problem}
associated to these strict inclusions (that is, given a function $f$, decide if
$f$ can be computed by a function of a proper subclass). Note that in
\cref{computational-models:fig}, if $\mathsf{B}$ is a proper subclass of
$\mathsf{C}$, then the \emph{aperiodic} variant of $\mathsf{B}$ is precisely
the intersection of $\mathsf{B}$ with the \emph{aperiodic} variant of the
larger class $\mathsf{C}$. As an example, to conclude that a function can be
computed by an \emph{aperiodic sequential function} ($\ASequential$) it is
enough to prove that it is sequential ($\Sequential$), and computed by an
aperiodic polyregular function ($\SF$).

\AP To approach the \emph{aperiodicity} membership problems the traditional
approach has been to develop so-called canonical objects, akin to the minimal
automaton or the syntactic monoid of regular languages, over which aperiodicity
is equivalent to a simple syntactic criterion (e.g. having no counters). This
is the proof scheme that has been implemented for sequential functions
\cite{CHOF03}, and more recently for rational functions \cite{FGL16,FGLM18}.
Unfortunately, such a canonical object remains to be defined for regular and
polyregular functions.

\begin{figure}
    \centering
    \begin{tikzpicture}[
        xscale=1.2,
        decidable/.style={
            ->, thick, color=black
        },
        unknown/.style={
            ->, dashed, color=gray
        }
        ]
        % Create nodes for the classes and their aperiodic variants
        % Mealy / sequential / rational / regular / polyregular
        \node (mealy)  at (0, 2) {$\Mealy$};
        \node (seq)    at (2, 2) {$\Sequential$};
        \node (rat)    at (4, 2) {$\Rational$};
        \node (reg)    at (6, 2) {$\Regular$};
        \node (poly)   at (8, 2) {$\Poly$};
        % Aperiodic variants
        \node (amealy) at (0,0) {$\AMealy$};
        \node (aseq)   at (2,0) {$\ASequential$};
        \node (arat)   at (4,0) {$\ARational$};
        \node (areg)   at (6,0) {$\ARegular$};
        \node (apoly)  at (8,0) {$\SF$};
        % Arrows
        % The decidability is known for inclusions
        % mealy -> sequential -> rational -> regular -> polyregular
        % similarly for the aperiodic variants
        \draw[decidable] (mealy) --
            node[midway, above] {folklore}
            (seq);
        \draw[decidable] (seq)   -- 
            node[midway, above] {\cite{CHOF03}}
            (rat);
        \draw[decidable] (rat)   -- 
            node[midway, above] {\cite{FGRS13,BGMP15}}
            (reg);
        \draw[decidable] (reg)   -- 
            node[midway, above] {\cite{BOKL19}}
            (poly);
        \draw[decidable] (amealy) --
            (aseq);
        \draw[decidable] (aseq) --
            (arat);
        \draw[decidable] (arat) --
            (areg);
        \draw[decidable] (areg) --
            (apoly);
        % Decidability of the aperidoic variants are known
        % for mealy, sequential, rational but not for regular nor polyregular.
        % we draw the arrows
        \draw[decidable] (amealy) --
            node[midway, left] {\cite{SCHU65,MNPA71}}
            (mealy);
        \draw[decidable] (aseq) -- 
            node[midway, left] {\cite{CHOF03}}
            (seq);
        \draw[decidable] (arat) -- 
            node[midway, left] {\cite{FGL16,FGLM18}}
            (rat);
        \draw[unknown] (areg) --
            node[midway, left] {$\simeq$\cite{BOJA14}}
            (reg);
        \draw[unknown] (apoly) --
            node[midway, left] {?}
            (poly);
    \end{tikzpicture}
    \caption{
        In this picture, arrows denote strict inclusions. Plain
        arrows signify that the corresponding membership problem is decidable,
        while dashed arrows signify that the corresponding membership problem
        is conjectured to be decidable. In the case of \kl{regular functions}
        ($\Regular$),
        the decidability has only been obtained under a different notion
        of equivalence of functions \cite{BOJA14}.
    }
    \label{computational-models:fig}
\end{figure}

\AP In order to approach the decidability of decision problems on
\kl{polyregular functions}, restricting the output alphabet to a single letter
has proven to be a fruitful approach \cite{DOUE21,DOUE22}. The key advantage of
single-letter output alphabets is that the particular order in which letters
are produced by the machine is irrelevant. This allows to obtain results that
would otherwise fail because of this ordering, such as the so-called
\emph{pebble minimization theorem} of \cite{DOUE21}, which fails in the general
case \cite{BOJA22,KLEP23}. Let us remark that the restriction to \emph{unary
output} in the case of \kl{polyregular functions} can often be lifted to
outputs in the free commutative semigroup generated by a finite alphabet
$\Gamma$. In this sense, the shift to \emph{single letter output alphabet} is
really about enforcing commutativity on the output. However, even with the
restriction to single letter output alphabets, the decidability of
\kl{star-free polyregular functions} among \kl{polyregular functions} remains
an open problem \cite[Conjecture 7.61]{DOUE23}.

\AP \kl{Polyregular functions} with a single letter output alphabet can be seen
as having outputs in $\Nat$, and this is why they were introduced respectively
as \kl{$\Nat$-polyregular functions} ($\intro*\NPoly$) and \kl{star-free
$\Nat$-polyregular functions} ($\intro*\NSF$) \cite{DOUE21,DOUE22}. From this
point of view, $\NPoly$ is a subclass of a larger and well-studied class of
functions called (non-commutative) \intro{rational series}
\cite{REUT80,BERE88,BERE10}. It has been noticed in \cite{CDTL23} that
\kl{$\Nat$-polyregular functions} correspond to a natural subclass of
\kl{$\Nat$-rational series} that was frequently rediscovered in the literature
\cite{SCHU62,KRRC13}. In \cite{CDTL23}, the authors introduced the class of
\kl{$\Rel$-polyregular functions} ($\intro*\ZPoly$) that generalize
\kl{$\Nat$-polyregular functions} by allowing negative outputs. This remains a
subclass of \kl{rational series}, more precisely, of the class of
\kl{$\Rel$-rational series} ($\intro*\ZRat$). For this class of functions the
aperiodic variant $\intro*\ZSF$ has been shown to be decidable \cite{CDTL23}.
Beware that concurrent (non-equivalent) notions of aperiodicity exists in the
research area of rational series \cite{REUT80,DRGA19,CDTL23}, and that we only
refer to the one applying to rational series of polynomial growths introduced
in \cite{CDTL23}.

\AP This approach through rational series has several limitations. First,
$\ZPoly$ does not correspond to a restriction of \kl{polyregular functions},
because of the extra expressive power introduced by the negative outputs
\cite{CDTL23}. Furthermore, deciding the membership of $\NRat$ inside $\ZRat$
is an open problem (see e.g. \cite{KARH77}), that remains open even in the
restricted case of $\NPoly$ and $\ZPoly$ \cite[Open question 5.55]{DOUE23}. It
was conjectured that, analogously to classes appearing in
\cref{computational-models:fig}, the following equation would hold: $\NSF =
\NPoly \cap \ZSF$ \cite[Conjecture 7.61]{DOUE23}. This would imply the
decidability of aperiodicity for single letter output polyregular functions
since aperiodicity is decidable for $\ZSF$. We refer the reader to
\cref{previously-known-inclusions:fig} to get an overview of the different
classes of functions appearing in the paper.

\begin{figure}
    \centering
    \includestandalone[width=7cm]{tikz/class-inclusions}
    \caption{
        Decidability and inclusions of classes of functions,
        arranged along two axes. The first one is the complexity
        of the output alphabet ($\Nat$, $\Rel$, $\Sigma$). The second
        one is the allowed computational power
        (star-free polyregular functions, polyregular functions, 
        rational series).
        Arrows denote strict inclusions,
        and effectiveness (both in terms of decidability and of effective
        representation) is represented by thick double arrows. Inclusions that are
        suspected to be effective are represented using a dashed arrow together with a
        question mark.
    }
    \label{previously-known-inclusions:fig}
\end{figure}

\paragraph*{Contributions.} This paper focuses on \kl{$\Nat$-polyregular
functions} that are furthermore \kl{commutative}. Remark that this enforces a
commutativity property both on the input alphabet and on the output alphabet.
In this restricted setting, we answer positively to \cite[Conjecture
7.61]{DOUE23} (restated in \cref{zsf-nsf:conjecture}), and prove that the
membership problems between $\NSF$, $\ZSF$, $\NPoly$, and $\ZPoly$ are all
decidable, with effective conversion algorithms. This effectively shows that
all arrows in the upper left square of \cref{previously-known-inclusions:fig}
are decidable in the \kl{commutative} setting. Furthermore, we characterize
\kl{commutative} \kl{$\Nat$-polyregular function} as the class of
\kl{$\Rel$-polyregular functions} that are \kl{combinatorial} (defined page
\kpageref{combinatorial}), which is a semantic property akin to the one
\kl{ultimate polynomiality} that characterizes $\ZSF$ inside $\ZPoly$
\cite{CDTL23}. As a byproduct of our analysis, we provide a counter example to
the characterization of polynomials computed by \kl{$\Nat$-rational series}
given by \cite[Theorem 3.3, page 4; re-introduced as \cref{karh:thm}]{KARH77}.

Finally, we make a first step towards understanding the non-commutative case by
introducing a canonical model of computation for \kl{$\Nat$-polyregular
functions} built upon residuals, called the \kl{residual transducer}. This
residual transducer is obtained by shifting attention from the notion of
equivalence relations of finite index over $\Sigma^*$, used in traditional
automata theory and in the study of $\ZPoly$, to the notion of
\kl{well-quasi-orderings}. The residual transducer is effectively computable if
$f$ is \kl{commutative} and \kl{$\Nat$-polyregular}, and it is conjectured to
be computable for all functions in $\NPoly$.

The approach taken in this paper avoids the use of tools from the theory of
factorizations forests in finite monoids \cite{SIMO90}, which has been the
traditional method used to understand (commutative output) \kl{polyregular
functions} \cite{DOUE21,DOUE22,CDTL23}, and instead
focuses on intrinsic description of the functions themselves (via canonical
computational models such as the \kl{residual transducer} of
\cref{residual-transducer:def}, or explicit encoding/decoding into polynomials
via \cref{decompose-polynomial:lem}).

\paragraph*{Outline of the paper.} In \cref{preliminaries:sec}, we introduce
the multiple characterizations of \kl{$\Nat$-polyregular function} (resp.
$\Rel$-polyregular functions), show that one can decide if a function $f \in
\ZPoly$ is \kl{commutative} (\cref{decidable-commutative-poly:lemma}), and give
the basic definitions used on multivariate polynomials. In
\cref{polynomials:sec}, we focus on \emph{polynomials}, by introducing the
\cref{karh:thm} of \cite{KARH77}, providing a counter-example in
\cref{thm:counter-example}, and an alternative characterization in
\cref{corrected-version:thm}. In \cref{beyond-polynomials:sec}, we leverage the
results over polynomials to decide $\NPoly$ inside $\ZPoly$ in the
\kl{commutative} case (\cref{decidable-n-poly:thm}), and prove that
\cref{zsf-nsf:conjecture}
\cite[Conjecture 7.61]{DOUE23} 
holds for \kl{commutative} functions
(\cref{zsf-npoly-nsf:thm}). Finally, in \cref{beyond-commutative:sec}, we
introduce the notion of \kl{residual transducer}
(\cref{residual-transducer:def}), and prove in \cref{non-commutative-npoly:thm}
that it characterizes $\NPoly$ inside $\ZPoly$ through a notion of
\kl{well-quasi-ordered} collection of residuals.