### 2024-08-27 parts/correction.tex
integer binomial polynomials

%! TeX program = xelatex
%! lang = en-US
\subsection{The Corrected Theorem}
\label{sec:proof}

\AP
The counterexample provided by \cref{def:bad-polynomial} relies on the fact
that $\CoveredPoly$ is not stable under fixing indeterminates, while
\kl{$\Nat$-polyregular functions} are.
Indeed, the polynomial $\BadPoly$ introduced in
\cref{def:bad-polynomial} satisfies $\BadPoly(X,Y,1) = 3X^2 + 3Y^2 - 2XY$, which has a negative
coefficient for a \kl{maximal monomial}, namely $-2XY$.
In this section, we prove that closing
$\CoveredPoly$ under variable assignments is enough to recover from
\cref{karh:thm}.
We use the following notation to fix the value of some indeterminate, if
$P(X,Y)$ is a polynomial in $\Rel[X,Y]$, then $\intro*\restr{P(X,Y)}{X = 1}$ is
the polynomial $P(1,Y) \in \Rel[Y]$. More generally, if $\nu$ is a partial
function from $\vec{X}$ to $\Nat$, written $\nu \colon \vec{X} \topartial
	\Nat$, the restriction $\restr{P(\vec{X})}{\nu}$ is the polynomial with
indeterminates $\vec{Y} \defined \vec{X} - \dom(\nu)$ obtained by fixing the
variables of the domain of $\nu$.

\begin{definition}
	The class $\intro*\CorrectPoly[\vec{X}]$ is the collection of
	polynomials $P \in \Rel[\vec{X}]$ such that,
	for every partial function $\nu \colon \vec{X} \topartial \Nat$,
	every \kl{maximal monomial} of
	$\restr{P}{\nu}$ is \kl{non-negative}.
\end{definition}

First, let us remark that $\CorrectPoly \subseteq \CoveredPoly$, because
polynomials in $\CorrectPoly$ are \kl{non-negative}. We also remarked at the
beginning of this section that our counterexample $\BadPoly$ provided in
\cref{thm:counter-example} is not in $\CorrectPoly$. The rest of the
section is mainly concerned with proving the following corrected version of
\cref{karh:thm}.

\begin{theorem}[restate=corrected-version:thm,label=corrected-version:thm]
	Let $P \in \Rel[\vec{X}]$.
	The following are equivalent:
	\begin{enumerate}
		\item \label{corrected-1:item} $P \in \CorrectPoly$,
		\item \label{corrected-2:item} $P$ is \kl{represented} by a \kl{$\Nat$-rational series},
		\item \label{corrected-3:item} $P$ is \kl{represented} by a \kl{$\Nat$-polyregular function},
		\item \label{corrected-4:item} $P$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function},
	\end{enumerate}
	Furthermore, the properties are decidable, and conversions effective.
\end{theorem}

\subsubsection{From $\Nat$-polyregular functions to polynomials}

\AP Let us now prove that \kl{$\Nat$-rational polynomials} are in
$\CorrectPoly$. This follows from the correct implication in the statement of
\cref{karh:thm}, but we provide a self-contained proof using a refinement of
the classical combinatorial arguments for $\ZPoly$ \cite[Lemma 4.16]{CDTL23}
and $\NPoly$ \cite[Lemma 5.37]{DOUE23}. We take extra care to reprove in our
upcoming \cref{n-poly-combinatorics:lem} a finer statement that has two main
goals: first, it formulates the traditional \emph{pumping argument} as a
composition property using \kl{(noncommutative output) polynomial functions},
allowing us to use the same formalism in the upcoming
\cref{k-combinatorial:def,ultimately-polynomial:def} and highlight the role of
\kl{commutative} \kl{polyregular functions} in the broader study of
\kl{polyregular functions}; then we also give a precise shape of the
polynomials that arise from such \emph{pumping arguments}, which was lacking in
former similar statements where $\Rat[X_1, \dots, X_n]$ was used as an over
approximation.

\AP To address our first goal, let us define that a function $q$ is a
\intro{(noncommutative output) polynomial function} from $\Nat^p$ to $\Sigma^*$
whenever there exists words $\alpha_0, \dots, \alpha_p \in \Sigma^*$, and words
$u_1, \dots, u_p \in \Sigma^*$, such that $q(X_1, \dots, X_p) = \alpha_0
	\prod_{i = 1}^p u_i^{X_i} \alpha_i$. That is, $q$ is syntactically defined by a
non-commutative monomial over the monoid $\Sigma^*$. Almost by
definition, \kl{polynomial functions} are \kl{commutative} \kl{polyregular
	functions}.

\AP Remark that our second goal is at odds with our restriction to polynomials
with coefficients in $\Rel$, because of the polynomial $X(X-1)/2$
(\cref{q-polynomial-n-poly:ex}). However, we argue that up to a suitable change
of basis for the ring of polynomials, all coefficients will actually be in
$\Nat$. This relatively simple trick works by considering as ``basic
polynomials" the \intro{binomial function} of the form $\binom{X - \ell}{k}$,
where $\ell$ and $k$ are natural numbers. We extend this to \intro{natural
	binomial functions} that are obtained by considering $\Nat$-linear combinations
of products of \kl{binomial functions}, that is, we consider functions that
have the following shape: $f(x_1, \dots, x_k) = \sum_{i = 1}^n n_i \prod_{j =
		1}^k \binom{x_j - p_{i,j}}{k_{i,j}}$. Beware that $\binom{X - \ell}{k}$ is
defined to be $0$ when $X \leq \ell$, and is therefore not a polynomial.

\begin{lemma}
	\label{binomial-function-star-free:lem}
	Let $F$ be a \kl{binomial function} from $\Nat^k$ to $\Nat$.
	Then it is \kl{represented} by a \kl{star-free polyregular function}.
\end{lemma}
\begin{proof}
	Because of the stability properties of
	$\NSF$ (\cref{stability-polyregular:lemma}), we only need to
	check that given $r,s \in \Nat$,
	the function $x \mapsto \binom{x - r}{s}$ is \kl{represented}
	by a function $f_{r,s} \in \NSF$.
	Let us prove it when $r = 0$, since the other functions
	can be obtained by translating $f_{r,s}$.
	Remark that
	$\binom{x}{s} = \sum_{u_1 \cdots u_{s+1} = a^x} 1$,
	which is clearly expressible as a \kl{star-free polyregular function}.
\end{proof}

\begin{lemma}
	\label{n-poly-combinatorics:lem}
	Let $f$ be an \kl{$\Nat$-polyregular} function.
	There exists a computable $\omega \in \Nat_{\geq 1}$
	such that for all \kl{polynomial functions}
	$q \colon \Nat^p \to \Sigma^*$,
	there exists a computable \kl{natural binomial function} $F$
	such that:
	\begin{equation*}
		f \circ q(\omega X_1, \dots, \omega X_p)
		=
		F
		\quad
		\text{ over } (\Nat_{\geq 1})^p
		\quad .
	\end{equation*}
\end{lemma}
\begin{proof}
	Let us fix a finite monoid $M$, a morphism $\mu \colon \Sigma^* \to M$, and
	let $\omega$ be an idempotent power for the finite monoid $M$,
	i.e., a number such that for all $x \in M$,
	$(x^{\omega})^2 = x^{\omega}$.
	We will prove by induction on $k \in \Nat_{\geq 1}$
	that for all
	$\pi \colon M^k \to \Nat$,
	the lemma holds with $s = 0$ for the function
	$f = \pi^\dagger$, i.e., such that
	$f(w)$ is the sum over all factorizations of $w$
	into $k$ words $(w_1, \dots, w_k) \in \Sigma^*$
	of the value $\pi(\mu(w_1), \dots, \mu(w_k))$.
	Since every \kl{$\Nat$-polyregular function}
	is obtained via some choice of $M, \mu, \pi$ (\cref{nat-rel-poly:def})
	we derive the desired result.

	In all cases, we will use a \kl{polynomial function} $q \colon \Nat^p \to
		\Sigma^*$ defined by $q(X_1, \dots, X_p) \defined \alpha_0 \prod_{i = 1}^p
		(u_i^{\omega \times X_i} \alpha_i)$ using some words $\alpha_i$ (for $0 \leq i
		\leq p$) and $u_i$ (for $1 \leq i \leq p)$ in $\Sigma^*$. That is, we will
	implicitly multiply all variables by $\omega$ in order to make the
	equations more readable. Note that when using an induction hypothesis, we
	will therefore have to check that the polynomial $q$ will have this
	specific form.

	\textbf{Base case: $k = 1$.}
	Let us show that $f \circ q$ is
	actually a constant (positive) function (and in particular a \kl{natural binomial function}), as for
	all $n_1, \dots, n_p \geq 1$,
	\begin{align*}
		f(q(n_1, \dots, n_p))
		 & \eqdef \pi\left(\mu(q(n_1, \dots, n_p))\right)                                      \\
		 & \eqdef \pi\left(\mu(\alpha_0 \prod_{i = 1}^p u_i^{\omega n_i} \alpha_i)\right)
		\\
		 & = \pi\left(\mu(\alpha_0) \prod_{i = 1}^p \mu(u_i)^{\omega n_i} \mu(\alpha_i)\right)
		 & \text{ morphism }
		\\
		 & = \pi\left(\mu(\alpha_0) \prod_{i = 1}^p \mu(u_i) \mu(\alpha_i)\right)
		 & \text{ idempotent and } n_i \geq 1                                                  \\
		 & = c \geq 0
	\end{align*}

	\textbf{Induction hypothesis.}
	The proof will essentially follow the same schema as the base case,
	with extra care needed to properly handle the summation of polynomials
	that arises when $k \geq 2$.

	We will use the fact that given $n_1, \dots, n_p \geq 0$,
	there exists a bijection between
	partitions of $q(\omega n_1, \dots, \omega n_p)$
	into $k+1$ words, and
	triples composed of
	a position  $1 \leq \ell \leq p$,
	a ratio     $0 \leq b \leq n_\ell$,
	a remainder $0 \leq r < \omega \times \card{u_\ell}$,
	and a partition of
	$)u_\ell^\omega)_{[-r:]} u_\ell^{\omega \times (n_\ell - b)}
		\alpha_{\ell+1} \cdots \alpha_{p-1} u_p^{n_p} \alpha_{p}$
	into $k$ words,
	where we have used the notation
	$(v)_{[-r:]}$ to denote the word obtained from $v$
	by keeping only the last $r$ letters, in a sort of Pythonic syntax.
	That is, we look at where the first ``cut" is made.

	Now, given $(\ell, b, r)$ respecting the inequalities
	mentioned above,
	one can define the \kl{polynomial functions}
	$s_{\ell,r,b} \colon \Nat^{\ell - 1} \to \Sigma^*$
	and
	$q_{\ell,r} \colon \Nat^{p - \ell + 1} \to \Sigma^*$
	so that
	$q(X_1, \dots, X_p) = s_{\ell,r,b}(X_1, \dots, X_{\ell-1})
		q_{\ell, r}(X_{\ell} - b, \dots, X_p)$
	under the assumption that $X_\ell \geq b$:
	\begin{equation*}
		q_{\ell,r}(X_{1}, \dots, X_{p - \ell + 1})
		\defined
		u_\ell^\omega\mid_{[-r:]} u_\ell^{\omega \times X_1}
		\alpha_{\ell+1} \cdots \alpha_{p-1} u_p^{X_{p - \ell + 1}} \alpha_{p}
		\quad .
	\end{equation*}
	Similarly, let us define
	$s_{\ell, r, b}(X_1, \dots, X_{\ell - 1})$
	as
	$\alpha_0 u_1^{\omega X_1} \alpha_2 \dots \alpha_{\ell - 1}
		u_\ell^{\omega b} (u_{\ell}^\omega)_{[1:r]}$,
	Remark that $q_{\ell,r}$ has all of its indeterminates
	multiplied by $\omega$, so that the induction hypothesis
	can be used.

	The key remark is that $\mu(s_{\ell,r,b}(X_1, \dots, X_{\ell - 1}))$ is a
	constant function from $\Nat^{\ell-1}$ to $M$, and that its value
	only depends on whether $b = 0$ or $b > 0$.
	This holds with
	the exact same proof as the base case of this induction. Let us call
	$m_{\ell,r,1}$ and $m_{\ell,r,0}$ the two elements of the monoid that are
	obtained by $\mu(s_{\ell,r,b}(X_1, \dots, X_{\ell-1}))$ respectively when
	$b = 0$ and when $b > 0$. Finally, let us define $\pi_{\ell,r,0} \colon M^k
		\to \Nat$ via $\pi_{\ell,r,0}(m_2, \dots, m_{k+1}) \defined
		\pi(m_{\ell,r,0}, m_1, \dots, m_{k+1})$.

	With all of this preliminary work done, we are now ready to conclude by
	simply unrolling the definition of
	a function
	$f$ defined using a \kl{production function}
	$\pi \colon M^{k+1} \to \Nat$.
	In the following equations, we omit the precise
	range for the sum over $(\ell, r, b)$ to save space.
	Similarly, we will use
	$q_{\ell,r}(X_1, \dots, X_{p-\ell+1})
		\defined q_{\ell,r}(X_1+b, X_2, \dots, X_{p - \ell + 1})$
	at the end of the computation.
	\begin{align*}
		 & f(q(n_1, \dots, n_p))           \\
		 & \eqdef
		\sum_{v_1 \dots v_{k+1} = q(n_1, \dots, n_p)}
		\pi(\mu(v_1), \dots, \mu(v_{k+1})) \\
		 & = \sum_{(\ell, b, r)}
		\sum_{v_2 \dots v_{k+1} = q_{\ell,r}(n_\ell - b, \dots, n_p)}
		\pi(\mu(s_{\ell,r,b}(n_1, \dots, n_{\ell-1})), \mu(v_2), \dots, \mu(v_{k+1}))
		\\
		 & =
		\sum_{(\ell, 0, r)}
		\sum_{v_2 \dots v_{k+1} = q_{\ell,b,0}(n_\ell - b, \dots, n_p)}
		\pi(m_{\ell,r,0}, \mu(v_2), \dots, \mu(v_{k+1}))
		\\
		 & +
		\sum_{(\ell, b > 0, r)}
		\sum_{v_2 \dots v_{k+1} = q_{\ell,r}(n_\ell - b, \dots, n_p)}
		\pi(m_{\ell,r,1}, \mu(v_2), \dots, \mu(v_{k+1}))
		\\
		 & =
		\sum_{(\ell, 0, r)}
		\sum_{v_2 \dots v_{k+1} = q_{\ell,b,0}(n_\ell - b, \dots, n_p)}
		\pi_{\ell,r,0}(\mu(v_2), \dots, \mu(v_{k+1}))
		\\
		 & +
		\sum_{(\ell, b > 0, r)}
		\sum_{v_2 \dots v_{k+1} = q_{\ell,r}(n_\ell-b, \dots, n_p)}
		\pi_{\ell,r,1}(\mu(v_2), \dots, \mu(v_{k+1}))
		\\
		 & =
		\sum_{1 \leq \ell \leq p}
		\sum_{0 \leq r < \omega \times \card{u_\ell}}
		(\pi_{\ell,r,0})^\dagger (q_{\ell, 0, r}(n_\ell-b, \dots, n_p))
		\\
		 & +
		\sum_{1 \leq \ell \leq p}
		\sum_{0 \leq r < \omega \times \card{u_\ell}}
		\sum_{0 < b < n_\ell}
		(\pi_{\ell,r,1})^{\dagger}(q_{\ell,r}(n_\ell - b, \dots, n_p))
	\end{align*}
	Now, using the induction hypothesis,
	we have \kl{natural binomial functions} $F_{\ell,r,0}$ and $F_{\ell,r,1}$
	satisfying
	\begin{equation*}
		(\pi_{\ell,r,1})^{\dagger}(q_{\ell,r}(n_\ell - b, \dots, n_p))
		= F_{\ell,r,1}(\omega (n_\ell - b), \dots, \omega n_p)
		\quad
		\forall n_\ell - b, \dots, n_p \geq 1
		\quad .
	\end{equation*}
	And similarly
	\begin{equation*}
		(\pi_{\ell,r,0})^{\dagger}(q_{\ell,r}(n_1, \dots, n_{\ell-1})
		= F_{\ell,r,0}(\omega n_1, \dots, \omega n_{\ell-1})
		\quad
		\forall n_1, \dots, n_{\ell - 1} \geq 1
		\quad .
	\end{equation*}

	It is a well-known equality that for all $s,n \in \Nat$,
	$\sum_{0 < i < n} \binom{s}{i} = \binom{s+1}{n}$.
	As a consequence, \kl{natural binomial functions} are
	stable under the following summation operation:
	$\Sigma_Y \colon Q(x_1,\dots,x_n,y) \mapsto \sum_{0 < b < y} Q(x_1, \dots, x_n, b)$.
	This allows us to conclude by defining:
	\begin{equation*}
		F(x_1, \dots, x_p) \defined
		\sum_{1 \leq \ell \leq p} \sum_{0 \leq r < \omega \card{u_\ell}}
		F_{l,r,0}(x_{\ell}, \dots, \omega x_p) + \Sigma_Y F_{l,r,1}(Y,
		x_{\ell+1}, \dots, x_p)
		\quad .
	\end{equation*}
	The latter being a \kl{natural binomial function} satisfying
	the desired equality.
\end{proof}

Let us highlight that the multiplicative factor $\omega$ is necessary in
\cref{n-poly-combinatorics:lem}. Indeed, the function $f \colon \set{a}^* \to
	\Nat$ defined as $0$ when the input is of odd length and $1$ when the input is
of even length is \kl{$\Nat$-polyregular}, but $f(a^X)$ is not a polynomial. We
can trade off this multiplicative factor for a constant term addition under the
extra assumption that the function is \kl{star-free polyregular}, as described
in the following \cref{n-sf-combinatorics:lem}.

\begin{lemma}
	\label{n-sf-combinatorics:lem}
	Let $f$ be a \kl{star-free $\Nat$-polyregular} function.
	There exists a computable $s \in \Nat_{\geq 1}$
	such that for all \kl{polynomial functions}
	$q \colon \Nat^p \to \Sigma^*$,
	there exists a computable \kl{natural binomial function} $F$
	such that:
	\begin{equation*}
		f \circ q(X_1 + s, \dots, X_p + s)
		=
		F
		\quad
		\text{ over } (\Nat_{\geq 1})^p
		\quad .
	\end{equation*}
\end{lemma}
\begin{proof}[Proof Sketch]
	We redo the proof of \cref{n-poly-combinatorics:lem}, with the extra
	assumption that the monoid $M$ is \kl(monoid){aperiodic}, hence that there
	exists $s \in \Nat$ such that $x^{s+1} = x^s$ for all $x \in M$. As a
	consequence, we can define $\omega = 1$, and
	translate inputs by $s$, instead of multiplying them by $\omega$.
\end{proof}

\begin{corollary}
	\label{n-rat-correct:lem}
	Let $P \in \Rel[X_1, \dots, X_p]$ be an \kl{$\Nat$-rational polynomial}.
	Then,
	$P \in \CorrectPoly$.
\end{corollary}
\begin{proof}
	Let $f$ be the \kl{commutative}
	\kl{$\Nat$-polyregular function}
	with domain defined as $\Sigma \defined \set{a_1, \dots, a_p}$
	that \kl{represents} $P$.
	Using \cref{n-poly-combinatorics:lem},
	there exists a number $\omega \in \Nat_{\geq 1}$
	and \kl{natural binomial function} $Q$
	such that
	for all $n_1, \dots, n_p \geq 1$:
	\begin{equation*}
		f\left(
		\prod_{i = 1}^p a_i^{\omega n_i}
		\right)
		= Q(n_1, \dots, n_p)
		= P(\omega n_1, \dots, \omega n_p)
		\quad .
	\end{equation*}
	Let us remark that for large enough values of $X$, the \kl{binomial function}
	$\binom{X - p}{k}$ coincides with a polynomial whose dominant coefficient
	is $1/k!$ which is \kl{non-negative}.
	We conclude that
	the \kl{maximal monomials} of
	$P(\omega X_1, \dots, \omega X_p)$ are \kl{non-negative},
	and since $\omega \geq 1$, we conclude that
	the \kl{maximal monomials} of $P$ have \kl{non-negative} coefficients.

	Remark that for every partial valuation $\nu \colon \vec{X} \topartial \Nat$,
	the polynomial $\restr{P}{\nu}$ continues to be represented
	by a \kl{$\Nat$-polyregular function}, namely
	$f$ partially applied to a word $u$ and using a restricted input alphabet. As a consequence,
	the \kl{maximal monomials} of
	$\restr{P}{\nu}$ are also \kl{non-negative},
	and
	we have proven that $P \in \CorrectPoly$.
\end{proof}

\subsubsection{From polynomials to $\Nat$-polyregular functions}

\AP We have proven that every \kl{$\Nat$-rational polynomial} is in
$\CorrectPoly$, and the rest of this section is devoted to proving the converse
implication. The core of the upcoming \cref{lem:correct-to-n-rat} leverages a
notion of \intro{discrete derivative} which is built by translating the domain
of the polynomial. To that end, let us write $\intro*\translate{K}$ for the
\intro{translation function} that maps a polynomial $P \in \Rel[X_1, \dots,
		X_n]$ to the polynomial $P(X_1 + K, \dots, X_n + K)$.

\begin{definition}
	\label{discrete-derivative:def}
	Let
	$K \in \Nat$,
	and
	$P \in \Rel[\vec{X}]$ be a polynomial,
	then
	%\begin{equation*}
	$
		\intro*\Diff{K}{P} \defined
		\translate{K}(P) - P
	$.
	%    \quad .
	%\end{equation*}
\end{definition}

\AP The key lemma of this section is \cref{lem:correct-to-n-rat}, which is
proved by induction on the number of indeterminates of a given polynomial $P$.
Notice that in \cref{lem:correct-to-n-rat}, polynomials can be represented
using \kl{star-free $\Nat$-polyregular functions}, which is a stronger
statement than the one in \cref{karh:thm}, but is in line with the idea that
polynomials are not periodic functions (this will be made precise in
\cref{zsf-nsf:conjecture}). In order to prove that result, we use the
combinatorial \cref{derivation-stabilises-correct:lem} that allows us to
transform a polynomial $P \in \CorrectPoly$ into a polynomial in
$\Nat[\vec{X}]$ through a well-chosen translation of the indeterminates.

\begin{lemma}
	\label{all-positive-derivative:lem}
	Let $P \in \Nat[\vec{X}]$ that is non-constant, and $K \in \Nat$,
	then $\Diff{K}{P} \in \Nat[\vec{X}]$ and all of its
	coefficients are (positive) multiples of $K$.
	Furthermore, every monomial that strictly divides some monomial of $P$
	appears in $\Diff{K}{P}$.
\end{lemma}
\begin{proof}
	We prove the result for monomials, as it extends
	to $\Nat$-linear combinations by linearity.
	Let $P = \prod_{i = 1}^k X_i^{\alpha_i}$ be a monomial.
	Notice that $\translate{K}(P) = \prod_{i = 1}^n (X_i + K)^{\alpha_i}$,
	and using a binomial expansion
	we list all the possible divisors of $P$,
	all of which with coefficients that are positive integers and multiples of $K$ except the coefficient
	of the maximal monomial (equal to $P$ itself) which is $1$.
	As a consequence, $\translate{K}(P) - P$ is simply
	obtained by removing this maximal monomial, which concludes the proof.
\end{proof}

\begin{lemma}
	\label{derivation-stabilises-correct:lem}
	Let $P \in \CorrectPoly$,
	$P_1$ be the sum of \kl{maximal monomials} of $P$,
	and $P_2 \defined P - P_1$ be the sum of
	non-maximal monomials of $P$.
	There exists a computable number $K \in \Nat$,
	such that
	$Q \defined (\Diff{K}{P_1} + \translate{K}(P_2)) \in \Nat[\vec{X}]$.
\end{lemma}
\begin{proof}
	Let us first tackle the specific case where $P$ is a constant polynomial.
	In this case, $P_1 = P$ and $P_2 = 0$.
	Furthermore,
	$\translate{K}(P_1) = P_1$ and $\Diff{K}{P_1} = 0$ for all $K \in \Nat$.
	We conclude that $\Diff{K}{P_1} + \translate{K}(P_2) = 0$
	for all $K \in \Nat$, hence belongs to $\Nat[\vec{X}]$. Selecting $K = 0$
	we conclude.

	Assume now that $P$ is not a constant polynomial. We will use
	\cref{all-positive-derivative:lem} on a well-selected value of $K$. Let us
	write $\alpha$ to be the maximal absolute value of a coefficient in $P$.
	Let $D$ be the number of unitary monomials that divide some monomial
	appearing in $P$. We can now define $K \defined D \times \alpha$,
	and let
	$Q \defined (\Diff{K}{P_1} + \translate{K}(P_2))$.
	Remark that $\Diff{K}{P_1}$ is already in $\Nat[\vec{X}]$,
	and the constant term of $\translate{K}(P_2)$ is also
	in $\Nat$.
	For any other monomial of $P_2$, by the maximality of $P_1$,
	it strictly divides some monomial of $P_1$, and
	equals some monomial of $\Diff{K}{P_1}$. Because every monomial
	of $\Diff{K}{P_1}$ has a coefficient that is a multiple of $K = \alpha \times D$, we can
	compensate every monomial of $P_2$ by a monomial of $\Diff{K}{P_1}$.
	Therefore,
	we conclude that
	$Q \in \Nat[\vec{X}]$.
\end{proof}

\begin{lemma}
	\label{lem:correct-to-n-rat}
	Let $P \in \Rel[\vec{X}]$.
	If $P \in \CorrectPoly$, then $P$ is \kl{represented}
	by a \kl{star-free $\Nat$-polyregular function},
	which is computable given $P$.
\end{lemma}
\begin{proof}
	We prove the result by induction on the number of indeterminates of $P$.
	In the proof, we write $\vec{X}$ for the indeterminates appearing in $P$,
	that is, we assume without loss of generality that all indeterminates are used.

	\textbf{Base case:} If the (unique) \kl{maximal monomial} of $P$ is a
	constant term. Since $P \in \CorrectPoly$, $P = n \in \Nat$, and therefore
	$P$ is \kl{represented} by a constant \kl{star-free $\Nat$-polyregular
		function}.

	\textbf{Induction:} Assume that $P$ is not a constant polynomial, and let
	us write $P = P_1 + P_2$ where $P_1$ is the sum of the \kl{maximal
		monomials} of $P$. We compute a bound $K$ such that $Q \defined
		(\Diff{K}{P_1} + \translate{K}(P_2)) \in \Nat[\vec{X}]$ using
	\cref{derivation-stabilises-correct:lem}. In particular, $Q$ is
	\kl{represented} by a \kl{star-free $\Nat$-polyregular function} using
	\cref{n-poly-n-poly:example}, the latter being effectively computable from
	$Q$. Let us now remark that $P_1 \in \Nat[\vec{X}]$, and is therefore
	(effectively) \kl{represented} by a \kl{star-free $\Nat$-polyregular
		function} (using again \cref{n-poly-n-poly:example}). As a consequence,
	$\translate{K}(P) = P_1 + Q$ is (effectively) \kl{represented} by a
	function $f_\Delta$.

	For all partial valuations $\nu \colon \vec{X} \topartial \set{0, \dots,
			K}$ fixing at least one indeterminate, one can use the induction hypothesis
	to compute a \kl{star-free $\Nat$-polyregular function} $f_\nu$ that
	\kl{represents} $\restr{P}{\nu}$. This is possible because we assumed that
	all indeterminates in $\vec{X}$ are used in $P$.

	Let us assume that the alphabet over which the (\kl{commutative}) functions
	$f_\Delta$ and $f_\nu$ are defined is $\set{a_1, \dots, a_n}$, with $a_i$
	representing the indeterminate $X_i$ of the polynomials. Now, let us define
	by case analysis the following \kl{commutative} \kl{star-free
		$\Nat$-polyregular function}, defined on words $w$ of the form $w \defined
		a_1^{x_1} \cdots a_n^{x_n}$, with $x_1, \dots, x_n \geq 0$.

	\begin{equation*}
		f(w) \defined
		\begin{cases}
			f_{[X_i \mapsto x_i]}(w)                     & \text{ if } \exists i \in \set{1, \dots, n}, x_i \leq K \\
			f_\Delta(a_1^{x_1 - K} \cdots a_n^{x_n - K}) & \text{ otherwise }
		\end{cases}
		\quad .
	\end{equation*}
	Remark that
	$f$ is a \kl{commutative} \kl{star-free $\Nat$-polyregular function}
	that
	\kl{represents} $P$.
\end{proof}

While \cref{lem:correct-to-n-rat} provides an effective conversion procedure,
it does not explicitly state that the membership is decidable to keep the proof
clearer. A similar proof scheme can be followed to conclude that membership is
decidable, and even show that elements in $\CorrectPoly$ are, up to suitable
translations, polynomials in $\Nat[\vec{X}]$. Beware that partial applications
are still needed in this characterization, as \cref{bad-poly-translate:ex}
illustrates. Effectiveness was the last missing piece of our main result
\cref{corrected-version:thm}, which corrects and generalizes \cref{karh:thm}.
\begin{example}
	\label{bad-poly-translate:ex}
	The polynomial $\BadPoly$ is not a
	\kl{$\Nat$-rational polynomial},
	but is \kl{non-negative} and satisfies
	$\translate{10}(\BadPoly) \in \Nat[\vec{X}]$.
\end{example}

\begin{lemma}
	\label{derivation-translation:lem}
	\proofref{derivation-translation:lem}
	Let $P \in \Rel[\vec{X}]$.
	There exists a computable number $K \in \Nat$
	such that the following are equivalent:
	\begin{enumerate}
		\item \label{d-t-correct:item} $P \in \CorrectPoly$,
		\item \label{d-t-transl:item}
		      for
		      all partial functions $\nu \colon \vec{X} \topartial \Nat$,
		      $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$,
		\item \label{d-t-transl-fin:item}
		      for all partial functions
		      $\nu \colon \vec{X} \topartial \set{0, \dots, K}$,
		      $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.
	\end{enumerate}
	In particular, the above properties are decidable.
\end{lemma}

\begin{proofof}[corrected-version:thm]
	The implications
	\cref{corrected-4:item} $\implies$
	\cref{corrected-3:item} $\implies$
	\cref{corrected-2:item} are obvious.
	\cref{lem:correct-to-n-rat} proves
	\cref{corrected-1:item} $\implies$ \cref{corrected-4:item},
	while \cref{n-rat-correct:lem}
	proves
	\cref{corrected-2:item} $\implies$ \cref{corrected-1:item}.
	Note that the lemmas provide effective conversion procedures,
	and that
	\cref{derivation-translation:lem}
	also provides a decision
	procedure.
\end{proofof}

For completeness, let us remark that the counterexample of
\cref{thm:counter-example} uses three indeterminates, and this is not a
coincidence: in the particular cases of one or two indeterminates,  the classes
$\CorrectPoly$ and $\CoveredPoly$ coincide. In particular, the examples
appearing in \cite{KARH77} are not invalidated, as they all use at most two
indeterminates. Note that the equivalence is clear for the univariate case,
where being non-negative and having non-negative maximal coefficient clearly
imply being an \kl{$\Nat$-rational polynomial}.

\begin{lemma}
	\label{lem:correct-covered-2}
	$\CorrectPoly[X,Y] = \CoveredPoly[X,Y]$.
\end{lemma}
\begin{proof}
	It is clear that $\CorrectPoly[X,Y] \subseteq \CoveredPoly[X,Y]$,
	by considering the empty valuation $\nu \colon \set{X,Y} \topartial \Nat$.
	For the converse inclusion, let us consider $P(X,Y)$
	that is \kl{non-negative}, such that the \kl{maximal monomials}
	are  \kl{non-negative}.

	If we fix none of the variables, then the \kl{maximal monomials}
	are \kl{non-negative} by assumption. If we fix one of the
	variables, we can assume without loss of generality that we
	fix $X = k$ for some $k \in \Nat$.
	Then $P(k,Y)$ is a \kl{non-negative} \emph{univariate} polynomial,
	and therefore must have a positive leading coefficient
	(which is the unique \kl{maximal monomial} in this case)
	or be constant equal to 0. In both cases, the \kl{maximal monomials}
	have positive coefficient.
	The same reasoning applies \emph{a fortiori} in the case where
	we fix the two indeterminates, leading to a constant polynomial.
\end{proof}

\subsubsection{From $\Rel$ to $\Rat$}

\AP Let us conclude this section by considering polynomials with
coefficients in $\Rat$, and justify that all the combinatorial work has
already happened in $\Rel$ and $\Nat$. A first observation is that from
\cref{n-poly-combinatorics:lem,n-sf-combinatorics:lem}, we know that the
functions that can be computed by \kl{star-free $\Nat$-polyregular
	functions} are going to be suitable combinations of \kl{natural binomial
	functions}. This motivates the following definition of the ``polynomial
counterpart" of a binomial coefficient, $\intro*\pbinom{X - \ell}{k}$
defined as $(X - \ell) \cdots (X - \ell - k) / k!$, that we call a
\intro{binomial monomial}, and introduce \intro{natural binomial
	polynomials} as $\Nat$-linear combinations of products of \kl{binomial
	monomials}, i.e., of the shape: $P(X_1, \dots, X_k) = \sum_{i = 1}^n n_i
	\prod_{j = 1}^k \pbinom{X_j - p_{i,j}}{k_{i,j}}$. Similarly, we introduce
the class of \intro{integer binomial polynomials}, which are obtained by
$\Rel$-linear combinations of products of \kl{binomial monomials}.

\AP
These definitions are motivated by the classical result of Pólya that
characterizes polynomials $P$ in $\Rat[X]$ that \intro{integer-valued} (i.e., are such
that $P(\Rel) \subseteq \Rel$) as \kl{integer binomial polynomials}
\cite{POLYA1915,CACHA1996}, which are often called \emph{Newton Polynomials}. Note
that this result straightforwardly extends to multiple indeterminates as we
prove in \cref{integer-binomial-polynomial:lem}.

\begin{lemma}
	\label{integer-binomial-polynomial:lem}
	Let $P \in \Rat[X_1, \dots, X_k]$ be a polynomial.
	Then, $P$ is an \kl{integer binomial polynomial} if and only if
	$P(\Rel^k) \subseteq \Rel$.
\end{lemma}
\begin{proof}
	Let us first notice that $\pbinom{X}{k}$ is \kl{integer valued}
	when $k \in \Nat$. One way to notice this is that when $X \geq 0$,
	$\pbinom{X}{k} = \binom{X}{k}$ is the number of ways to choose $k$ elements among $X$,
	which is an integer. When
	$X < 0$, then $\pbinom{X}{k} = (-1)^k \pbinom{k - X + 1}{k} = (-1)^k \binom{k - X + 1}{k}$
	is also an integer. Because \kl{integer-valued} polynomials are closed under products
	and $\Rel$-linear combination, we conclude that all \kl{integer binomial polynomials} are
	\kl{integer-valued}.

	For the converse direction, we follow the original proof of Pólya \cite{POLYA1915,CACHA1996}.
	The only difference is that we leverage differential operators on multiple indeterminates.
	To that end, let us consider the operator $\Delta_i$ that maps
	a polynomial $P$ to the polynomial $\Delta_i(P) \defined P(X_1, \dots, X_i + 1, \dots, X_k) - P$.
	That is, the discrete partial derivative with respect to the $i$-th indeterminate.
	Let us notice that $\Delta_i \pbinom{X_i}{\alpha} = \pbinom{X_i}{\alpha - 1}$ when $\alpha > 0$,
	indeed:
	\begin{align*}
		\Delta_i \pbinom{X_i}{\alpha} & = \pbinom{X_i + 1}{\alpha} - \pbinom{X_i}{\alpha}                                                         \\
		                              & = \frac{(X_i + 1) \cdots (X_i + 1 - \alpha + 1)}{\alpha!} - \frac{X_i \cdots (X_i - \alpha + 1)}{\alpha!} \\
		                              & = \frac{X_i \cdots (X_i - \alpha + 2)}{\alpha!} \alpha                                                    \\
		                              & = \pbinom{X_i}{\alpha - 1}
		\quad .
	\end{align*}
	Furthermore, when $\alpha = 0$, $\Delta_i \pbinom{X_i}{0} = 0$.

	Let us now remark that a simple expansion proves $\Delta_i \Delta_j P = \Delta_j \Delta_i P$ for all
	$1 \leq i,j \leq k$.
	This justifies the soundness of the notation
	$\Delta_\alpha$ where $\alpha \colon \set{1, \dots, k} \to \Nat$ is a
	multi-index, where $\Delta_{\alpha} \defined \Delta_1^{\alpha_1} \cdots
		\Delta_k^{\alpha_k}$.
	Let us write $\card{\alpha} \defined \sum_{i = 1}^k \alpha_i$ for the
	size of a multi-index.
	We claim that for all $P \in \Rat[X_1, \dots, X_k]$ of degree at most $d$,
	the following \emph{discrete Taylor expansion} holds:
	\begin{equation}
		\label{discrete-taylor:eq}
		P =
		\sum_{\card{\alpha} \leq d}
		\Delta_{\alpha} P (0, \dots, 0)
		\times
		\prod_{i = 1}^k \pbinom{X_i}{\alpha_i}
		\quad .
	\end{equation}

	To obtain \cref{discrete-taylor:eq}, we proceed by induction on the degree $d$.
	When $d = 0$, the result is holds because $P = P(0, \dots, 0)$.
	Assume that the result holds for all polynomials of degree at most $d$.
	Let $P$ of degree at most $d+1$.
	Then, there exists coefficients $\seqof{c_{\alpha}}{\card{\alpha} \leq d+1}$
	in $\Rat$
	such that
	\begin{equation*}
		P = \sum_{\card{\alpha} \leq d+1} c_{\alpha} \prod_{i = 1}^k \pbinom{X_i}{\alpha_i}
		\quad .
	\end{equation*}
	This is because the monomials of the form $\prod_{i = 1}^k \pbinom{X_i}{\alpha_i}$
	form a basis of $\Rat[X_1, \dots, X_k]$.
	In particular, when applying one partial derivative, one can leverage the induction hypothesis to obtain:
	\begin{align*}
		\Delta_1 P & = \sum_{\card{\alpha} \leq d+1} c_{\alpha} \Delta_1 \prod_{i = 1}^k \pbinom{X_i}{\alpha_i}
		=
		\sum_{\card{\alpha} \leq d+1 \wedge \alpha_1 > 0} c_{\alpha}  \pbinom{X_1}{\alpha_1 - 1} \prod_{i = 2}^k \pbinom{X_i}{\alpha_i}
		\\
		           & = \sum_{\card{\beta} \leq d} \Delta_{\beta}(\Delta_1 P)(0,\dots,0) \times \pbinom{X_1}{\beta_1} \prod_{i = 2}^k \pbinom{X_i}{\beta_i}
		\quad .
	\end{align*}

	Using again the fact that the monomials form a basis, we conclude by
	identification that for all $\card{\alpha} \leq d+1$ such that $\alpha_1 >
		0$, we have $c_{\alpha} = \Delta_{\alpha}P (0,\dots,0)$. We can repeat the
	same process with $\Delta_i$ for $1 \leq i \leq k$ to obtain that
	$c_{\alpha} = \Delta_{\alpha} P (0,\dots, 0)$ for all $\alpha$ such that $1
		\leq \card{\alpha} \leq d+1$. Now, for the specific case of $\card{\alpha}
		= 0$, which is the constant term of $P$, we simply conclude that
	$P(0,\dots,0) = c_{\alpha}$, and notice that $\Delta_{\alpha} P = P$ in
	this case.

	Let us now conclude by noticing that if $P$ is \kl{integer-valued},
	then so is $\Delta_i P$ for all $1 \leq i \leq k$. As a consequence,
	so are $\Delta_{\alpha} P$, for all multi-indexes $\alpha$, and we conclude
	that $\Delta_{\alpha}P (0,\dots,0) \in \Rel$ for all $\alpha$.
	Combined with \cref{discrete-taylor:eq}, we conclude that $P$ is an \kl{integer binomial polynomial}.
\end{proof}

As an immediate corollary, we completely characterize the class of
polynomials with rational coefficients that can be computed by
\kl{star-free $\Nat$-polyregular functions}.

\begin{corollary}
	\label{integer-binomial-polynomials:cor}
	Let $P \in \Rat[\vec{X}]$, and let $\alpha$ be the minimal number in $\Nat$ such that
	$\alpha P \in \Rel[\vec{X}]$. Then, the following properties are equivalent:
	\begin{enumerate}
		\item \label{int-bin-0:item} $P$ is an \kl{integer binomial polynomial},
		\item \label{int-bin-1:item} $P$ \kl{represented} by a \kl{star-free $\Rel$-polyregular function},
		\item \label{int-bin-2:item} $P$ \kl{represented} by a \kl{$\Rel$-polyregular function},
		\item \label{int-bin-3:item} $P(\Rel^k) \subseteq \Rel$.
	\end{enumerate}
	These properties are furthermore decidable.
\end{corollary}
\begin{proof}
	Let us first note that implications \cref{int-bin-1:item}
	$\implies$ \cref{int-bin-2:item} $\implies$ \cref{int-bin-3:item} are obvious.
	Now, \cref{int-bin-3:item} $\implies$ \cref{int-bin-0:item} is a direct
	consequence of \cref{integer-binomial-polynomial:lem}.
	Finally, \cref{int-bin-0:item} $\implies$ \cref{int-bin-1:item} follows from the fact
	that $\pbinom{X - p}{k}$ is \kl{represented} by a \kl{star-free $\Rel$-polyregular function}
	defined by hardcoding the output values (in $\Rel$) when $0 \leq X \leq p$, and
	using a \kl{star-free $\Nat$-polyregular function} when $X > p$
	(\cref{binomial-function-star-free:lem}).
	Because
	$\ZSF$ is closed under products and $\Rel$-linear combinations, we conclude.
\end{proof}

Obtaining an analogue of \cref{integer-binomial-polynomials:cor}
for \kl{$\Nat$-polyregular functions} is more challenging. The
naïve approach would be to consider some variation around \kl{natural binomial
	polynomials}. A first observation is that the class of products of the form
$\prod_{i = 1}^n \pbinom{X_i - p_i}{k_i}$ is a generating family for
$\Rat[\vec{X}]$ but is not linearly independent. As a consequence, there may be
several ways to write a given polynomial, and as illustrated by
\cref{natural-binomial-polynomial-basis:ex}, some of them will
exhibit negative coefficients.\footnote{This was not a problem for \kl{integer
		binomial polynomials}, for which the change of basis mapped coefficients in
	$\Rel$ to coefficients $\Rel$.} Furthermore, there are some \kl{natural
	binomial polynomials} that take negative values on positive inputs, as
illustrated by \cref{natural-binomial-polynomial-negative:ex}.
Finally, restricting our attention to \kl{non-negative} \kl{natural binomial
	polynomials} suffers from the same kind of issues as
\cref{karh:thm},
witnessed by \cref{natural-binomial-polynomial-positive-bad:ex}.

\begin{example}
	\label{natural-binomial-polynomial-negative:ex}
	The polynomial $\pbinom{X - 2}{2}$ takes negative values for $X = 0$.
\end{example}

\begin{example}
	\label{natural-binomial-polynomial-basis:ex}
	The polynomial $\pbinom{X - 2}{2} + 1$ is a \kl{non-negative} \kl{natural binomial polynomial}
	but in the standard basis, it is written
	$\pbinom{X}{2} - \pbinom{X}{1} + 2$, which has a negative coefficient.
\end{example}

\begin{example}
	\label{natural-binomial-polynomial-positive-bad:ex}
	The polynomial
	$Q(X,Y,Z) \defined \pbinom{X - 4}{1} \pbinom{Y}{1} \pbinom{Z}{1}
		+ 8 \pbinom{Y}{2} + 8 \pbinom{Z}{2} + 4$
	is a \kl{non-negative} \kl{natural binomial polynomial},
	but cannot be computed by a \kl{star-free $\Nat$-polyregular function}.
\end{example}

\AP We say that a \kl{natural binomial polynomial} is \intro{in base $K$} if
all its \kl{binomial monomials} are of the form $\pbinom{X - K}{d}$ for some $d
	\in \Nat$. To address the issues raised by the previous examples, we introduce
the notion of \intro{strongly natural binomial polynomials} \kl{in base $K$},
as the polynomials $P \in \Rat[X]$ such that for all partial valuation $\nu
	\colon \topartial \set{0, \dots, K}$, $\restr{P}{\nu}$ is a \kl{natural
	binomial polynomial} \kl{in base $K$}. This leads to the following
characterization.

\begin{corollary}
	\label{decide-rat-poly-npoly:cor}
	Let $P \in \Rat[\vec{X}]$ be a polynomial with \emph{rational}
	coefficients and let $\alpha$ be the smallest number in $\Nat_{\geq 1}$
	such that $\alpha P \in \Rel[\vec{X}]$. Then, there exists a
	computable $K \in \Nat$, such that the following are equivalent:
	\begin{enumerate}
		\item \label{rat-npoly-1:item}
		      $P$ is a \kl{strongly natural binomial polynomial} \kl{in base $K$},
		\item \label{rat-npoly-2:item} $P$ is \kl{represented} by a \kl{$\Nat$-polyregular function},
		\item \label{rat-npoly-3:item} $P$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function},
		\item \label{rat-npoly-4:item} $\alpha P \in \CorrectPoly$,
		      and for all $\vec{x} \in \Nat^{\vec{X}}$,
		      $P(\vec{x})$ is a multiple of $\alpha$.
	\end{enumerate}
	In particular, the properties are decidable.
\end{corollary}
\begin{proof}
	First, \cref{rat-npoly-4:item} $\implies$ \cref{rat-npoly-3:item}, because
	if $\alpha P \in \CorrectPoly$, then it is an \kl{$\Nat$-rational
		polynomial}. But in particular, there exists a polyregular function $f$
	that represents $\alpha P$ with outputs in $\set{ a }$. Now, let us
	post-compose this function by the one that maps $a^{k \alpha + r}$ to
	$a^{k}$, which is a \kl{polyregular function}. We conclude that $P$ is
	\kl{represented} by an \kl{$\Nat$-polyregular function}.

	We also know that \cref{rat-npoly-2:item} $\implies$
	\cref{rat-npoly-4:item}, since if $P$ is an \kl{$\Nat$-rational
		polynomial}, then so is $\alpha P$, and in particular $\alpha P \in
		\CorrectPoly$ by \cref{corrected-version:thm}.
\end{proof}