### 2024-09-02 parts/correction.tex

%! TeX program = xelatex
%! lang = en-US

\AP We will discuss at the end of \cref{sec:poly-to-n-poly}
why the
example of \cref{thm:counter-example} is minimal in the number of
indeterminates, which first requires us to provide a \emph{correct} analogue of
\cref{karh:thm}. The
counterexample provided by \cref{def:bad-polynomial}
relies on the fact that
$\CoveredPoly$ is not stable under fixing indeterminates, while
\kl{$\Nat$-rational polynomials} are. Indeed, the polynomial $\BadPoly$
introduced in \cref{def:bad-polynomial}
satisfies $\BadPoly(X,Y,1) = 3X^2 +
3Y^2 - 2XY$, which has a negative coefficient for a \kl{maximal monomial},
namely $-2XY$. In this section, we prove that closing $\CoveredPoly$ under
variable assignments is enough to recover from
\cref{karh:thm}. We
use the following notation to fix the value of some indeterminate, if $P(X,Y)$
is a polynomial in $\Rel[X,Y]$, then $\intro*\restr{P(X,Y)}{X = 1}$ is the
polynomial $P(1,Y) \in \Rel[Y]$. More generally, if $\nu$ is a partial function
from $\vec{X}$ to $\Nat$, written $\nu \colon \vec{X} \topartial \Nat$, the
restriction $\restr{P(\vec{X})}{\nu}$ is the polynomial with indeterminates
$\vec{Y} \defined \vec{X} - \dom(\nu)$ obtained by fixing the variables of the
domain of $\nu$.

\begin{definition}
	The class $\intro*\CorrectPoly[\vec{X}]$ is the collection of
	polynomials $P \in \Rel[\vec{X}]$ such that,
	for every partial function $\nu \colon \vec{X} \topartial \Nat$,
	every \kl{maximal monomial} of
	$\restr{P}{\nu}$ is \kl{non-negative}.
\end{definition}

First, let us remark that $\CorrectPoly \subseteq \CoveredPoly$, because
polynomials in $\CorrectPoly$ are \kl{non-negative}. We also remarked at the
beginning of this section that our counterexample $\BadPoly$ provided in
\cref{thm:counter-example} is not in $\CorrectPoly$. The rest of the
section is mainly concerned with proving the following corrected version of
\cref{karh:thm}.

\begin{theorem}[restate=corrected-version:thm,label=corrected-version:thm]
	Let $P \in \Rel[\vec{X}]$.
	The following are equivalent:
	\begin{enumerate}
		\item \label{corrected-1:item} $P \in \CorrectPoly$,
		\item \label{corrected-2:item} $P$ is \kl{represented} by a \kl{$\Nat$-rational series},
		\item \label{corrected-3:item} $P$ is \kl{represented} by a \kl{$\Nat$-polyregular function},
		\item \label{corrected-4:item} $P$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function},
	\end{enumerate}
	Furthermore, the properties are decidable, and conversions effective.
\end{theorem}

Let us briefly remark that \cref{corrected-version:thm} is surprising given the
fact that is not possible to decide whether a polynomial $P \in \Rel[\vec{X}]$
is \kl{non-negative} (\cref{undecidable-non-negative:lem}), by
reduction to the undecidability of Hilbert's Tenth Problem
\cite{HILB1902,MATI1970}.

\begin{lemma}
    \label{undecidable-non-negative:lem}
    Let $P \in \Rel[\vec{X}]$ be a polynomial. Then, checking whether $P$ is
    \kl{non-negative} is undecidable.
\end{lemma}
\begin{proof}
    Recall that by the undecidability of the Hilbert's Tenth Problem \cite[Problem 10 page 22]{HILB1902}, it is
    undecidable whether a given polynomial $P \in \Rel[\vec{X}]$ has a solution in
    $\Rel^n$ \cite{MATI1970,DAVIS1973}.
    The result follows by observing that
    checking whether a polynomial $P \in \Rel[\vec{X}]$ has no solutions in $\Rel^n$ if and only 
    if $P^2 - 1$ is \kl{non-negative}. 
\end{proof}

The proof of \cref{corrected-version:thm} is divided into two parts. First, we
provide in \cref{sec:n-poly-to-poly}
a fine
combinatorial understanding of what functions can be computed by functions in
$\NPoly$ and $\ZPoly$. This will allow us to prove that \kl{$\Nat$-polyregular
functions} are in $\CorrectPoly$ (\cref{n-rat-correct:lem}).
Then, in \cref{sec:poly-to-n-poly}
we will
show how to compute polynomials in $\CorrectPoly$ using $\NPoly$
(\cref{lem:correct-to-n-rat}).
With this in hand, we will next generalize these results
to polynomials in $\Rat[\vec{X}]$ in \cref{sec:rel-to-rat}.

\subsection{From $\Nat$-polyregular functions to polynomials}
\label{sec:n-poly-to-poly}

\AP Let us prove that \kl{$\Nat$-rational polynomials} are in $\CorrectPoly$.
This fact follows from the correct implication in the statement of
\cref{karh:thm}, but
we provide a self-contained proof using a refinement of the classical
combinatorial \emph{pumping arguments} for $\ZPoly$ \cite[Lemma 4.16]{CDTL23} and
$\NPoly$ \cite[Lemma 5.37]{DOUE23}. We take extra care to reprove in our
upcoming \cref{n-poly-combinatorics:lem} a strong statement that has
two main goals. Our first goal is to highlight the role of \kl{commutative}
\kl{polyregular functions} in the broader study of \kl{polyregular functions},
which is done by reformulating the traditional pumping argument as a
composition property involving said functions, which will be reused in the
upcoming \cref{k-combinatorial:def,ultimately-polynomial:def}
in
\cref{beyond-polynomials:sec}. Our second goal
is to give a precise shape of the functions that arise from such \emph{pumping
arguments}, which was lacking in former similar statements.

\AP To address our first goal, let us define that a function $q$ is a
\intro{(noncommutative output) polynomial function} from $\Nat^p$ to $\Sigma^*$
whenever there exists words $\alpha_0, \dots, \alpha_p \in \Sigma^*$, and words
$u_1, \dots, u_p \in \Sigma^*$, such that $q(X_1, \dots, X_p) = \alpha_0
	\prod_{i = 1}^p u_i^{X_i} \alpha_i$. That is, $q$ is syntactically defined by a
non-commutative monomial over the monoid $\Sigma^*$. Almost by
definition, \kl{polynomial functions} are \kl{commutative} \kl{polyregular
	functions}.

\AP Our second goal is achieved by understanding that \kl{$\Nat$-polyregular
functions} essentially compute binomial coefficients, as illustrated by the
polynomial $X(X-1)/2 = \binom{X}{2}$.
of \cref{q-polynomial-n-poly:ex}. A \intro{binomial function} is a
function of the form $\binom{X - \ell}{k}$, where $\ell$ and $k$ are natural
numbers. We extend this to \intro{natural binomial functions} that are obtained
by considering $\Nat$-linear combinations of products of \kl{binomial
functions}, that is, we consider functions that have the following shape:
$f(x_1, \dots, x_k) = \sum_{i = 1}^n n_i \prod_{j = 1}^k \binom{x_j -
p_{i,j}}{k_{i,j}}$. Beware that $\binom{X - \ell}{k}$ is defined to be $0$ when
$X \leq \ell$, and is therefore not a polynomial. Let us immediately
prove that \kl{binomial functions}  can be \kl{represented} in $\NSF$,
generalizing \cref{q-polynomial-n-poly:ex}.

\begin{lemma}
	\label{binomial-function-star-free:lem}
	Let $F$ be a \kl{binomial function} from $\Nat^k$ to $\Nat$.
	Then it is \kl{represented} by a \kl{star-free polyregular function}.
\end{lemma}
\begin{proof}
	Because of the stability properties of
	$\NSF$ (\cref{stability-polyregular:lemma}), we only need to
	check that given $r,s \in \Nat$,
	the function $x \mapsto \binom{x - r}{s}$ is \kl{represented}
	by a function $f_{r,s} \in \NSF$.
	Let us prove it when $r = 0$, since the other functions
	can be obtained by translating $f_{r,s}$.
	Remark that
	$\binom{x}{s} = \sum_{u_1 \cdots u_{s+1} = a^x} 1$,
	which is clearly expressible as a \kl{star-free polyregular function}.
\end{proof}

\begin{lemma}
	\label{n-poly-combinatorics:lem}
	Let $f$ be an \kl{$\Nat$-polyregular} function.
	There exists a computable $\omega \in \Nat_{\geq 1}$
	such that for all \kl{polynomial functions}
	$q \colon \Nat^p \to \Sigma^*$,
	there exists a computable \kl{natural binomial function} $F$
	such that:
	\begin{equation*}
		f \circ q(\omega X_1, \dots, \omega X_p)
		=
		F
		\quad
		\text{ over } (\Nat_{\geq 1})^p
		\quad .
	\end{equation*}
\end{lemma}
\begin{proof}
	Let us fix a finite monoid $M$, a morphism $\mu \colon \Sigma^* \to M$, and
	let $\omega$ be an idempotent power for the finite monoid $M$,
	i.e., a number such that for all $x \in M$,
	$(x^{\omega})^2 = x^{\omega}$.
	We will prove by induction on $k \in \Nat_{\geq 1}$
	that for all
	$\pi \colon M^k \to \Nat$,
	the lemma holds with $s = 0$ for the function
	$f = \pi^\dagger$, i.e., such that
	$f(w)$ is the sum over all factorizations of $w$
	into $k$ words $(w_1, \dots, w_k) \in \Sigma^*$
	of the value $\pi(\mu(w_1), \dots, \mu(w_k))$.
	Since every \kl{$\Nat$-polyregular function}
	is obtained via some choice of $M, \mu, \pi$ (\cref{nat-rel-poly:def})
	we derive the desired result.

	In all cases, we will use a \kl{polynomial function} $q \colon \Nat^p \to
		\Sigma^*$ defined by $q(X_1, \dots, X_p) \defined \alpha_0 \prod_{i = 1}^p
		(u_i^{\omega \times X_i} \alpha_i)$ using some words $\alpha_i$ (for $0 \leq i
		\leq p$) and $u_i$ (for $1 \leq i \leq p)$ in $\Sigma^*$. That is, we will
	implicitly multiply all variables by $\omega$ in order to make the
	equations more readable. Note that when using an induction hypothesis, we
	will therefore have to check that the polynomial $q$ will have this
	specific form.

	\textbf{Base case: $k = 1$.}
	Let us show that $f \circ q$ is
	actually a constant (positive) function (and in particular a \kl{natural binomial function}), as for
	all $n_1, \dots, n_p \geq 1$,
	\begin{align*}
		f(q(n_1, \dots, n_p))
		 & \eqdef \pi\left(\mu(q(n_1, \dots, n_p))\right)                                      \\
		 & \eqdef \pi\left(\mu(\alpha_0 \prod_{i = 1}^p u_i^{\omega n_i} \alpha_i)\right)
		\\
		 & = \pi\left(\mu(\alpha_0) \prod_{i = 1}^p \mu(u_i)^{\omega n_i} \mu(\alpha_i)\right)
		 & \text{ morphism }
		\\
		 & = \pi\left(\mu(\alpha_0) \prod_{i = 1}^p \mu(u_i) \mu(\alpha_i)\right)
		 & \text{ idempotent and } n_i \geq 1                                                  \\
		 & = c \geq 0
	\end{align*}

	\textbf{Induction hypothesis.}
	The proof will essentially follow the same schema as the base case,
	with extra care needed to properly handle the summation of polynomials
	that arises when $k \geq 2$.

	We will use the fact that given $n_1, \dots, n_p \geq 0$,
	there exists a bijection between
	partitions of $q(\omega n_1, \dots, \omega n_p)$
	into $k+1$ words, and
	triples composed of
	a position  $1 \leq \ell \leq p$,
	a ratio     $0 \leq b \leq n_\ell$,
	a remainder $0 \leq r < \omega \times \card{u_\ell}$,
	and a partition of
	$)u_\ell^\omega)_{[-r:]} u_\ell^{\omega \times (n_\ell - b)}
		\alpha_{\ell+1} \cdots \alpha_{p-1} u_p^{n_p} \alpha_{p}$
	into $k$ words,
	where we have used the notation
	$(v)_{[-r:]}$ to denote the word obtained from $v$
	by keeping only the last $r$ letters, in a sort of Pythonic syntax.
	That is, we look at where the first ``cut" is made.

	Now, given $(\ell, b, r)$ respecting the inequalities
	mentioned above,
	one can define the \kl{polynomial functions}
	$s_{\ell,r,b} \colon \Nat^{\ell - 1} \to \Sigma^*$
	and
	$q_{\ell,r} \colon \Nat^{p - \ell + 1} \to \Sigma^*$
	so that
	$q(X_1, \dots, X_p) = s_{\ell,r,b}(X_1, \dots, X_{\ell-1})
		q_{\ell, r}(X_{\ell} - b, \dots, X_p)$
	under the assumption that $X_\ell \geq b$:
	\begin{equation*}
		q_{\ell,r}(X_{1}, \dots, X_{p - \ell + 1})
		\defined
		u_\ell^\omega\mid_{[-r:]} u_\ell^{\omega \times X_1}
		\alpha_{\ell+1} \cdots \alpha_{p-1} u_p^{X_{p - \ell + 1}} \alpha_{p}
		\quad .
	\end{equation*}
	Similarly, let us define
	$s_{\ell, r, b}(X_1, \dots, X_{\ell - 1})$
	as
	$\alpha_0 u_1^{\omega X_1} \alpha_2 \dots \alpha_{\ell - 1}
		u_\ell^{\omega b} (u_{\ell}^\omega)_{[1:r]}$,
	Remark that $q_{\ell,r}$ has all of its indeterminates
	multiplied by $\omega$, so that the induction hypothesis
	can be used.

	The key remark is that $\mu(s_{\ell,r,b}(X_1, \dots, X_{\ell - 1}))$ is a
	constant function from $\Nat^{\ell-1}$ to $M$, and that its value
	only depends on whether $b = 0$ or $b > 0$.
	This holds with
	the exact same proof as the base case of this induction. Let us call
	$m_{\ell,r,1}$ and $m_{\ell,r,0}$ the two elements of the monoid that are
	obtained by $\mu(s_{\ell,r,b}(X_1, \dots, X_{\ell-1}))$ respectively when
	$b = 0$ and when $b > 0$. Finally, let us define $\pi_{\ell,r,0} \colon M^k
		\to \Nat$ via $\pi_{\ell,r,0}(m_2, \dots, m_{k+1}) \defined
		\pi(m_{\ell,r,0}, m_1, \dots, m_{k+1})$.

	With all of this preliminary work done, we are now ready to conclude by
	simply unrolling the definition of
	a function
	$f$ defined using a \kl{production function}
	$\pi \colon M^{k+1} \to \Nat$.
	In the following equations, we omit the precise
	range for the sum over $(\ell, r, b)$ to save space.
	Similarly, we will use
	$q_{\ell,r}(X_1, \dots, X_{p-\ell+1})
		\defined q_{\ell,r}(X_1+b, X_2, \dots, X_{p - \ell + 1})$
	at the end of the computation.
	\begin{align*}
		 & f(q(n_1, \dots, n_p))           \\
		 & \eqdef
		\sum_{v_1 \dots v_{k+1} = q(n_1, \dots, n_p)}
		\pi(\mu(v_1), \dots, \mu(v_{k+1})) \\
		 & = \sum_{(\ell, b, r)}
		\sum_{v_2 \dots v_{k+1} = q_{\ell,r}(n_\ell - b, \dots, n_p)}
		\pi(\mu(s_{\ell,r,b}(n_1, \dots, n_{\ell-1})), \mu(v_2), \dots, \mu(v_{k+1}))
		\\
		 & =
		\sum_{(\ell, 0, r)}
		\sum_{v_2 \dots v_{k+1} = q_{\ell,b,0}(n_\ell - b, \dots, n_p)}
		\pi(m_{\ell,r,0}, \mu(v_2), \dots, \mu(v_{k+1}))
		\\
		 & +
		\sum_{(\ell, b > 0, r)}
		\sum_{v_2 \dots v_{k+1} = q_{\ell,r}(n_\ell - b, \dots, n_p)}
		\pi(m_{\ell,r,1}, \mu(v_2), \dots, \mu(v_{k+1}))
		\\
		 & =
		\sum_{(\ell, 0, r)}
		\sum_{v_2 \dots v_{k+1} = q_{\ell,b,0}(n_\ell - b, \dots, n_p)}
		\pi_{\ell,r,0}(\mu(v_2), \dots, \mu(v_{k+1}))
		\\
		 & +
		\sum_{(\ell, b > 0, r)}
		\sum_{v_2 \dots v_{k+1} = q_{\ell,r}(n_\ell-b, \dots, n_p)}
		\pi_{\ell,r,1}(\mu(v_2), \dots, \mu(v_{k+1}))
		\\
		 & =
		\sum_{1 \leq \ell \leq p}
		\sum_{0 \leq r < \omega \times \card{u_\ell}}
		(\pi_{\ell,r,0})^\dagger (q_{\ell, 0, r}(n_\ell-b, \dots, n_p))
		\\
		 & +
		\sum_{1 \leq \ell \leq p}
		\sum_{0 \leq r < \omega \times \card{u_\ell}}
		\sum_{0 < b < n_\ell}
		(\pi_{\ell,r,1})^{\dagger}(q_{\ell,r}(n_\ell - b, \dots, n_p))
	\end{align*}
	Now, using the induction hypothesis,
	we have \kl{natural binomial functions} $F_{\ell,r,0}$ and $F_{\ell,r,1}$
	satisfying
	\begin{equation*}
		(\pi_{\ell,r,1})^{\dagger}(q_{\ell,r}(n_\ell - b, \dots, n_p))
		= F_{\ell,r,1}(\omega (n_\ell - b), \dots, \omega n_p)
		\quad
		\forall n_\ell - b, \dots, n_p \geq 1
		\quad .
	\end{equation*}
	And similarly
	\begin{equation*}
		(\pi_{\ell,r,0})^{\dagger}(q_{\ell,r}(n_1, \dots, n_{\ell-1})
		= F_{\ell,r,0}(\omega n_1, \dots, \omega n_{\ell-1})
		\quad
		\forall n_1, \dots, n_{\ell - 1} \geq 1
		\quad .
	\end{equation*}

	It is a well-known equality that for all $s,n \in \Nat$,
	$\sum_{0 < i < n} \binom{s}{i} = \binom{s+1}{n}$.
	As a consequence, \kl{natural binomial functions} are
	stable under the following summation operation:
	$\Sigma_Y \colon Q(x_1,\dots,x_n,y) \mapsto \sum_{0 < b < y} Q(x_1, \dots, x_n, b)$.
	This allows us to conclude by defining:
	\begin{equation*}
		F(x_1, \dots, x_p) \defined
		\sum_{1 \leq \ell \leq p} \sum_{0 \leq r < \omega \card{u_\ell}}
		F_{l,r,0}(x_{\ell}, \dots, \omega x_p) + \Sigma_Y F_{l,r,1}(Y,
		x_{\ell+1}, \dots, x_p)
		\quad .
	\end{equation*}
	The latter being a \kl{natural binomial function} satisfying
	the desired equality.
\end{proof}

Let us highlight that the multiplicative factor $\omega$ is necessary in
\cref{n-poly-combinatorics:lem}. Indeed, the function $f \colon \set{a}^* \to
	\Nat$ defined as $0$ when the input is of odd length and $1$ when the input is
of even length is \kl{$\Nat$-polyregular}, but $f(a^X)$ is not a polynomial. We
can trade off this multiplicative factor for a constant term addition under the
extra assumption that the function is \kl{star-free polyregular}, as described
in the following \cref{n-sf-combinatorics:lem}.

\begin{lemma}
	\label{n-sf-combinatorics:lem}
	Let $f$ be a \kl{star-free $\Nat$-polyregular} function.
	There exists a computable $s \in \Nat_{\geq 1}$
	such that for all \kl{polynomial functions}
	$q \colon \Nat^p \to \Sigma^*$,
	there exists a computable \kl{natural binomial function} $F$
	such that:
	\begin{equation*}
		f \circ q(X_1 + s, \dots, X_p + s)
		=
		F
		\quad
		\text{ over } \Nat^p
		\quad .
	\end{equation*}
\end{lemma}
\begin{proof}[Proof Sketch]
	We redo the proof of \cref{n-poly-combinatorics:lem}, with the extra
	assumption that the monoid $M$ is \kl(monoid){aperiodic}, hence that there
	exists $s \in \Nat$ such that $x^{s+1} = x^s$ for all $x \in M$. As a
	consequence, we can define $\omega = 1$, and
	translate inputs by $s$, instead of multiplying them by $\omega$.
\end{proof}

Because \kl{natural binomial functions} behave as polynomials with
\kl{non-negative} \kl{maximal monomials} on large enough inputs, we can
conclude from \cref{n-poly-combinatorics:lem}
that \kl{$\Nat$-rational polynomials} are in $\CorrectPoly$.

\begin{corollary}
	\label{n-rat-correct:lem}
	Let $P \in \Rel[X_1, \dots, X_p]$ be an \kl{$\Nat$-rational polynomial}.
	Then,
	$P \in \CorrectPoly$.
\end{corollary}
\begin{proof}
	Let $f$ be the \kl{commutative}
	\kl{$\Nat$-polyregular function}
	with domain defined as $\Sigma \defined \set{a_1, \dots, a_p}$
	that \kl{represents} $P$.
	Using \cref{n-poly-combinatorics:lem},
	there exists a number $\omega \in \Nat_{\geq 1}$
	and \kl{natural binomial function} $Q$
	such that
	for all $n_1, \dots, n_p \geq 1$:
	\begin{equation*}
		f\left(
		\prod_{i = 1}^p a_i^{\omega n_i}
		\right)
		= Q(n_1, \dots, n_p)
		= P(\omega n_1, \dots, \omega n_p)
		\quad .
	\end{equation*}
	Let us remark that for large enough values of $X$, the \kl{binomial function}
	$\binom{X - p}{k}$ coincides with a polynomial whose dominant coefficient
	is $1/k!$ which is \kl{non-negative}.
	We conclude that
	the \kl{maximal monomials} of
	$P(\omega X_1, \dots, \omega X_p)$ are \kl{non-negative},
	and since $\omega \geq 1$, we conclude that
	the \kl{maximal monomials} of $P$ have \kl{non-negative} coefficients.

	Remark that for every partial valuation $\nu \colon \vec{X} \topartial \Nat$,
	the polynomial $\restr{P}{\nu}$ continues to be represented
	by a \kl{$\Nat$-polyregular function}, namely
	$f$ partially applied to a word $u$ and using a restricted input alphabet. As a consequence,
	the \kl{maximal monomials} of
	$\restr{P}{\nu}$ are also \kl{non-negative},
	and
	we have proven that $P \in \CorrectPoly$.
\end{proof}

\subsection{From polynomials to $\Nat$-polyregular functions}
\label{sec:poly-to-n-poly}

\AP We have proven that every \kl{$\Nat$-rational polynomial} is in
$\CorrectPoly$, and the rest of this section is devoted to proving the converse
implication. The core of the upcoming \cref{lem:correct-to-n-rat}
leverages a notion of \intro{discrete derivative} which is built by
translating the domain of the polynomial. To that end, let us write
$\intro*\translate{K}$ for the \intro{translation function} that maps a
polynomial $P \in \Rel[X_1, \dots, X_n]$ to the polynomial $P(X_1 + K, \dots,
X_n + K)$.

\begin{definition}
	\label{discrete-derivative:def}
	Let
	$K \in \Nat$,
	and
	$P \in \Rel[\vec{X}]$ be a polynomial,
	then
	%\begin{equation*}
	$
		\intro*\Diff{K}{P} \defined
		\translate{K}(P) - P
	$.
	%    \quad .
	%\end{equation*}
\end{definition}

\AP The key lemma of this section is \cref{lem:correct-to-n-rat}, which is
proved by induction on the number of indeterminates of a given polynomial $P$.
Notice that in \cref{lem:correct-to-n-rat}, polynomials can be represented
using \kl{star-free $\Nat$-polyregular functions}, which is a stronger
statement than the one in \cref{karh:thm}, but is in line with the idea that
polynomials are not periodic functions (this will be made precise in
\cref{zsf-nsf:conjecture}). In order to prove that result, we use the
combinatorial \cref{derivation-stabilises-correct:lem} that allows us to
transform a polynomial $P \in \CorrectPoly$ into a polynomial in
$\Nat[\vec{X}]$ through a well-chosen translation of the indeterminates.

\begin{lemma}
	\label{all-positive-derivative:lem}
	Let $P \in \Nat[\vec{X}]$ that is non-constant, and $K \in \Nat$,
	then $\Diff{K}{P} \in \Nat[\vec{X}]$ and all of its
	coefficients are (positive) multiples of $K$.
	Furthermore, every monomial that strictly divides some monomial of $P$
	appears in $\Diff{K}{P}$.
\end{lemma}
\begin{proof}
	We prove the result for monomials, as it extends
	to $\Nat$-linear combinations by linearity.
	Let $P = \prod_{i = 1}^k X_i^{\alpha_i}$ be a monomial.
	Notice that $\translate{K}(P) = \prod_{i = 1}^n (X_i + K)^{\alpha_i}$,
	and using a binomial expansion
	we list all the possible divisors of $P$,
	all of which with coefficients that are positive integers and multiples of $K$ except the coefficient
	of the maximal monomial (equal to $P$ itself) which is $1$.
	As a consequence, $\translate{K}(P) - P$ is simply
	obtained by removing this maximal monomial, which concludes the proof.
\end{proof}

\begin{lemma}
	\label{derivation-stabilises-correct:lem}
	Let $P \in \CorrectPoly$,
	$P_1$ be the sum of \kl{maximal monomials} of $P$,
	and $P_2 \defined P - P_1$ be the sum of
	non-maximal monomials of $P$.
	There exists a computable number $K \in \Nat$,
	such that
	$Q \defined (\Diff{K}{P_1} + \translate{K}(P_2)) \in \Nat[\vec{X}]$.
\end{lemma}
\begin{proof}
	Let us first tackle the specific case where $P$ is a constant polynomial.
	In this case, $P_1 = P$ and $P_2 = 0$.
	Furthermore,
	$\translate{K}(P_1) = P_1$ and $\Diff{K}{P_1} = 0$ for all $K \in \Nat$.
	We conclude that $\Diff{K}{P_1} + \translate{K}(P_2) = 0$
	for all $K \in \Nat$, hence belongs to $\Nat[\vec{X}]$. Selecting $K = 0$
	we conclude.

	Assume now that $P$ is not a constant polynomial. We will use
	\cref{all-positive-derivative:lem} on a well-selected value of $K$. Let us
	write $\alpha$ to be the maximal absolute value of a coefficient in $P$.
	Let $D$ be the number of unitary monomials that divide some monomial
	appearing in $P$. We can now define $K \defined D \times \alpha$,
	and let
	$Q \defined (\Diff{K}{P_1} + \translate{K}(P_2))$.
	Remark that $\Diff{K}{P_1}$ is already in $\Nat[\vec{X}]$,
	and the constant term of $\translate{K}(P_2)$ is also
	in $\Nat$.
	For any other monomial of $P_2$, by the maximality of $P_1$,
	it strictly divides some monomial of $P_1$, and
	equals some monomial of $\Diff{K}{P_1}$. Because every monomial
	of $\Diff{K}{P_1}$ has a coefficient that is a multiple of $K = \alpha \times D$, we can
	compensate every monomial of $P_2$ by a monomial of $\Diff{K}{P_1}$.
	Therefore,
	we conclude that
	$Q \in \Nat[\vec{X}]$.
\end{proof}

\begin{lemma}
	\label{lem:correct-to-n-rat}
	Let $P \in \Rel[\vec{X}]$.
	If $P \in \CorrectPoly$, then $P$ is \kl{represented}
	by a \kl{star-free $\Nat$-polyregular function},
	which is computable given $P$.
\end{lemma}
\begin{proof}
	We prove the result by induction on the number of indeterminates of $P$.
	In the proof, we write $\vec{X}$ for the indeterminates appearing in $P$,
	that is, we assume without loss of generality that all indeterminates are used.

	\textbf{Base case:} If the (unique) \kl{maximal monomial} of $P$ is a
	constant term. Since $P \in \CorrectPoly$, $P = n \in \Nat$, and therefore
	$P$ is \kl{represented} by a constant \kl{star-free $\Nat$-polyregular
		function}.

	\textbf{Induction:} Assume that $P$ is not a constant polynomial, and let
	us write $P = P_1 + P_2$ where $P_1$ is the sum of the \kl{maximal
		monomials} of $P$. We compute a bound $K$ such that $Q \defined
		(\Diff{K}{P_1} + \translate{K}(P_2)) \in \Nat[\vec{X}]$ using
	\cref{derivation-stabilises-correct:lem}. In particular, $Q$ is
	\kl{represented} by a \kl{star-free $\Nat$-polyregular function} using
	\cref{n-poly-n-poly:example}, the latter being effectively computable from
	$Q$. Let us now remark that $P_1 \in \Nat[\vec{X}]$, and is therefore
	(effectively) \kl{represented} by a \kl{star-free $\Nat$-polyregular
		function} (using again \cref{n-poly-n-poly:example}). As a consequence,
	$\translate{K}(P) = P_1 + Q$ is (effectively) \kl{represented} by a
	function $f_\Delta$.

	For all partial valuations $\nu \colon \vec{X} \topartial \set{0, \dots,
			K}$ fixing at least one indeterminate, one can use the induction hypothesis
	to compute a \kl{star-free $\Nat$-polyregular function} $f_\nu$ that
	\kl{represents} $\restr{P}{\nu}$. This is possible because we assumed that
	all indeterminates in $\vec{X}$ are used in $P$.

	Let us assume that the alphabet over which the (\kl{commutative}) functions
	$f_\Delta$ and $f_\nu$ are defined is $\set{a_1, \dots, a_n}$, with $a_i$
	representing the indeterminate $X_i$ of the polynomials. Now, let us define
	by case analysis the following \kl{commutative} \kl{star-free
		$\Nat$-polyregular function}, defined on words $w$ of the form $w \defined
		a_1^{x_1} \cdots a_n^{x_n}$, with $x_1, \dots, x_n \geq 0$.

	\begin{equation*}
		f(w) \defined
		\begin{cases}
			f_{[X_i \mapsto x_i]}(w)                     & \text{ if } \exists i \in \set{1, \dots, n}, x_i \leq K \\
			f_\Delta(a_1^{x_1 - K} \cdots a_n^{x_n - K}) & \text{ otherwise }
		\end{cases}
		\quad .
	\end{equation*}
	Remark that
	$f$ is a \kl{commutative} \kl{star-free $\Nat$-polyregular function}
	that
	\kl{represents} $P$.
\end{proof}

While \cref{lem:correct-to-n-rat} provides an effective conversion procedure,
it does not explicitly state that the membership is decidable to keep the proof
clearer. A similar proof scheme can be followed to conclude that membership is
decidable, and even show that elements in $\CorrectPoly$ are, up to suitable
translations, polynomials in $\Nat[\vec{X}]$. Beware that partial applications
are still needed in this characterization, as \cref{bad-poly-translate:ex}
illustrates. Effectiveness was the last missing piece of our main result
\cref{corrected-version:thm}, which corrects and generalizes \cref{karh:thm}.
\begin{example}
	\label{bad-poly-translate:ex}
	The polynomial $\BadPoly$ is not a
	\kl{$\Nat$-rational polynomial},
	but is \kl{non-negative} and satisfies
	$\translate{10}(\BadPoly) \in \Nat[\vec{X}]$.
\end{example}

\begin{lemma}
	\label{derivation-translation:lem}
	\proofref{derivation-translation:lem}
	Let $P \in \Rel[\vec{X}]$.
	There exists a computable number $K \in \Nat$
	such that the following are equivalent:
	\begin{enumerate}
		\item \label{d-t-correct:item} $P \in \CorrectPoly$,
		\item \label{d-t-transl:item}
		      for
		      all partial functions $\nu \colon \vec{X} \topartial \Nat$,
		      $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$,
		\item \label{d-t-transl-fin:item}
		      for all partial functions
		      $\nu \colon \vec{X} \topartial \set{0, \dots, K}$,
		      $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.
	\end{enumerate}
	In particular, the above properties are decidable.
\end{lemma}

\begin{proofof}[corrected-version:thm]
	The implications
	\cref{corrected-4:item} $\implies$
	\cref{corrected-3:item} $\implies$
	\cref{corrected-2:item} are obvious.
	\cref{lem:correct-to-n-rat} proves
	\cref{corrected-1:item} $\implies$ \cref{corrected-4:item},
	while \cref{n-rat-correct:lem}
	proves
	\cref{corrected-2:item} $\implies$ \cref{corrected-1:item}.
	Note that the lemmas provide effective conversion procedures,
	and that
	\cref{derivation-translation:lem}
	also provides a decision
	procedure.
\end{proofof}

For completeness, let us remark that the counterexample of
\cref{thm:counter-example} uses three indeterminates, and this is not a
coincidence: in the particular cases of one or two indeterminates,  the classes
$\CorrectPoly$ and $\CoveredPoly$ coincide. In particular, the examples
appearing in \cite{KARH77} are not invalidated, as they all use at most two
indeterminates. Note that the equivalence is clear for the univariate case,
where being non-negative and having non-negative maximal coefficient clearly
imply being an \kl{$\Nat$-rational polynomial}.

\begin{lemma}
	\label{lem:correct-covered-2}
	$\CorrectPoly[X,Y] = \CoveredPoly[X,Y]$.
\end{lemma}
\begin{proof}
	It is clear that $\CorrectPoly[X,Y] \subseteq \CoveredPoly[X,Y]$,
	by considering the empty valuation $\nu \colon \set{X,Y} \topartial \Nat$.
	For the converse inclusion, let us consider $P(X,Y)$
	that is \kl{non-negative}, such that the \kl{maximal monomials}
	are  \kl{non-negative}.

	If we fix none of the variables, then the \kl{maximal monomials}
	are \kl{non-negative} by assumption. If we fix one of the
	variables, we can assume without loss of generality that we
	fix $X = k$ for some $k \in \Nat$.
	Then $P(k,Y)$ is a \kl{non-negative} \emph{univariate} polynomial,
	and therefore must have a positive leading coefficient
	(which is the unique \kl{maximal monomial} in this case)
	or be constant equal to 0. In both cases, the \kl{maximal monomials}
	have positive coefficient.
	The same reasoning applies \emph{a fortiori} in the case where
	we fix the two indeterminates, leading to a constant polynomial.
\end{proof}

\subsection{From $\Rel$ to $\Rat$}
\label{sec:rel-to-rat}

\AP Let us complete our analysis of polynomials \kl{represented} by $\NPoly$ or
$\ZPoly$ by considering polynomials with coefficients in $\Rat$, and justify
that all the combinatorial work has already happened in $\Rel$ and $\Nat$. A
first observation is that from \cref{n-sf-combinatorics:lem}, we
know that the polynomials that can be computed by \kl{star-free
$\Nat$-polyregular functions} are going to be suitable combinations of
\kl{natural binomial functions}. This motivates the following definition of the
``polynomial counterpart" of a binomial coefficient, given two numbers $\ell,k
\in \Nat$, $\intro*\pbinom{X - \ell}{k}$ defined as $(X - \ell) \cdots (X -
\ell - k) / k!$,\footnote{ In particular, $\pbinom{X - \ell}{k}$ is defined to
be $1$ when $k = 0$, and $X - \ell$ when $k = 1$. } that we call a
\intro{binomial monomial}, and introduce \intro{natural binomial polynomials}
as $\Nat$-linear combinations of products of \kl{binomial monomials}, i.e., of
the shape: $P(X_1, \dots, X_k) = \sum_{i = 1}^n n_i \prod_{j = 1}^k \pbinom{X_j
- p_{i,j}}{k_{i,j}}$. Similarly, we introduce the class of \intro{integer
binomial polynomials}, which are obtained by $\Rel$-linear combinations of
products of \kl{binomial monomials}.

\AP
These definitions are motivated by the classical result of Pólya that
characterizes polynomials $P$ in $\Rat[X]$ that \intro{integer-valued} (i.e., are such
that $P(\Rel) \subseteq \Rel$) as \kl{integer binomial polynomials}
\cite{POLYA1915,CACHA1996}, which are often called \emph{Newton Polynomials}. Note
that this result straightforwardly extends to multiple indeterminates as we
prove in \cref{integer-binomial-polynomial:lem}.

\begin{lemma}
	\label{integer-binomial-polynomial:lem}
	Let $P \in \Rat[X_1, \dots, X_k]$ be a polynomial.
	Then, $P$ is an \kl{integer binomial polynomial} if and only if
	$P(\Rel^k) \subseteq \Rel$, if and only if $P(\Nat^k) \subseteq \Rel$.
\end{lemma}
\begin{proof}
	Let us first notice that $\pbinom{X}{k}$ is \kl{integer-valued}
	when $k \in \Nat$. One way to notice this is that when $X \geq 0$,
	$\pbinom{X}{k} = \binom{X}{k}$ is the number of ways to choose $k$ elements among $X$,
	which is an integer. When
	$X < 0$, then $\pbinom{X}{k} = (-1)^k \pbinom{k - X + 1}{k} = (-1)^k \binom{k - X + 1}{k}$
	is also an integer. Because \kl{integer-valued} polynomials are closed under products,
	$\Rel$-linear combination, and 
    shifting of domain by an integer,
    we conclude that all \kl{integer binomial polynomials} are
	\kl{integer-valued}.

	For the converse direction, we follow the original proof of Pólya \cite{POLYA1915,CACHA1996}.
	The only difference is that we leverage differential operators on multiple indeterminates.
	To that end, let us consider the operator $\Delta_i$ that maps
	a polynomial $P$ to the polynomial $\Delta_i(P) \defined P(X_1, \dots, X_i + 1, \dots, X_k) - P$.
	That is, the discrete partial derivative with respect to the $i$-th indeterminate.
	Let us notice that $\Delta_i \pbinom{X_i}{\alpha} = \pbinom{X_i}{\alpha - 1}$ when $\alpha > 0$,
	indeed:
	\begin{align*}
		\Delta_i \pbinom{X_i}{\alpha} & = \pbinom{X_i + 1}{\alpha} - \pbinom{X_i}{\alpha}                                                         \\
		                              & = \frac{(X_i + 1) \cdots (X_i + 1 - \alpha + 1)}{\alpha!} - \frac{X_i \cdots (X_i - \alpha + 1)}{\alpha!} \\
		                              & = \frac{X_i \cdots (X_i - \alpha + 2)}{\alpha!} \alpha                                                    \\
		                              & = \pbinom{X_i}{\alpha - 1}
		\quad .
	\end{align*}
	Furthermore, when $\alpha = 0$, $\Delta_i \pbinom{X_i}{0} = 0$.

	Let us now remark that a simple expansion proves $\Delta_i \Delta_j P = \Delta_j \Delta_i P$ for all
	$1 \leq i,j \leq k$.
	This justifies the soundness of the notation
	$\Delta_\alpha$ where $\alpha \colon \set{1, \dots, k} \to \Nat$ is a
	multi-index, where $\Delta_{\alpha} \defined \Delta_1^{\alpha_1} \cdots
		\Delta_k^{\alpha_k}$.
	Let us write $\card{\alpha} \defined \sum_{i = 1}^k \alpha_i$ for the
	size of a multi-index.
	We claim that for all $P \in \Rat[X_1, \dots, X_k]$ of degree at most $d$,
	the following \emph{discrete Taylor expansion} holds:
	\begin{equation}
		\label{discrete-taylor:eq}
		P =
		\sum_{\card{\alpha} \leq d}
		\Delta_{\alpha} P (0, \dots, 0)
		\times
		\prod_{i = 1}^k \pbinom{X_i}{\alpha_i}
		\quad .
	\end{equation}

	To obtain \cref{discrete-taylor:eq}, we proceed by induction on the degree $d$.
	When $d = 0$, the result is holds because $P = P(0, \dots, 0)$.
	Assume that the result holds for all polynomials of degree at most $d$.
	Let $P$ of degree at most $d+1$.
	Then, there exists coefficients $\seqof{c_{\alpha}}{\card{\alpha} \leq d+1}$
	in $\Rat$
	such that
	\begin{equation*}
		P = \sum_{\card{\alpha} \leq d+1} c_{\alpha} \prod_{i = 1}^k \pbinom{X_i}{\alpha_i}
		\quad .
	\end{equation*}
	This is because the monomials of the form $\prod_{i = 1}^k \pbinom{X_i}{\alpha_i}$
	form a basis of $\Rat[X_1, \dots, X_k]$.
	In particular, when applying one partial derivative, one can leverage the induction hypothesis to obtain:
	\begin{align*}
		\Delta_1 P & = \sum_{\card{\alpha} \leq d+1} c_{\alpha} \Delta_1 \prod_{i = 1}^k \pbinom{X_i}{\alpha_i}
		=
		\sum_{\card{\alpha} \leq d+1 \wedge \alpha_1 > 0} c_{\alpha}  \pbinom{X_1}{\alpha_1 - 1} \prod_{i = 2}^k \pbinom{X_i}{\alpha_i}
		\\
		           & = \sum_{\card{\beta} \leq d} \Delta_{\beta}(\Delta_1 P)(0,\dots,0) \times \pbinom{X_1}{\beta_1} \prod_{i = 2}^k \pbinom{X_i}{\beta_i}
		\quad .
	\end{align*}

	Using again the fact that the monomials form a basis, we conclude by
	identification that for all $\card{\alpha} \leq d+1$ such that $\alpha_1 >
		0$, we have $c_{\alpha} = \Delta_{\alpha}P (0,\dots,0)$. We can repeat the
	same process with $\Delta_i$ for $1 \leq i \leq k$ to obtain that
	$c_{\alpha} = \Delta_{\alpha} P (0,\dots, 0)$ for all $\alpha$ such that $1
		\leq \card{\alpha} \leq d+1$. Now, for the specific case of $\card{\alpha}
		= 0$, which is the constant term of $P$, we simply conclude that
	$P(0,\dots,0) = c_{\alpha}$, and notice that $\Delta_{\alpha} P = P$ in
	this case.

	Let us now conclude by noticing that if $P$ is \kl{integer-valued},
	then so is $\Delta_i P$ for all $1 \leq i \leq k$. As a consequence,
	so are $\Delta_{\alpha} P$, for all multi-indexes $\alpha$, and we conclude
	that $\Delta_{\alpha}P (0,\dots,0) \in \Rel$ for all $\alpha$.
	Combined with \cref{discrete-taylor:eq}, we conclude that $P$ is an \kl{integer binomial polynomial}.
    Remark that $\Delta_{\alpha} P (0, \dots, 0)$ only ever involves evaluation of $P$ on 
    $\Nat^k$, justifying the weakening of the hypothesis.
\end{proof}

As an immediate corollary,  we completely characterize the class of polynomials
in $\Rat[\vec{X}]$ that are \kl{represented} by $\ZPoly$ as the \kl{integer
binomial polynomials}.

\begin{theorem}
	\label{integer-binomial-polynomials:cor}
	Let $P \in \Rat[\vec{X}]$, and let $\alpha$ be the minimal number in $\Nat$ such that
	$\alpha P \in \Rel[\vec{X}]$. Then, the following properties are equivalent:
	\begin{enumerate}
		\item \label{int-bin-0:item} $P$ is an \kl{integer binomial polynomial},
		\item \label{int-bin-1:item} $P$ \kl{represented} by a \kl{star-free $\Rel$-polyregular function},
		\item \label{int-bin-2:item} $P$ \kl{represented} by a \kl{$\Rel$-polyregular function},
        \item \label{int-bin-3:item} $P$ is \kl{integer-valued}.
	\end{enumerate}
	These properties are furthermore decidable.
\end{theorem}
\begin{proof}
	Let us first note that implications \cref{int-bin-1:item}
	$\implies$ \cref{int-bin-2:item} $\implies$ \cref{int-bin-3:item} are obvious.
	Now, \cref{int-bin-3:item} $\implies$ \cref{int-bin-0:item} is a direct
	consequence of \cref{integer-binomial-polynomial:lem}.
	Finally, \cref{int-bin-0:item} $\implies$ \cref{int-bin-1:item} follows from the fact
	that $\pbinom{X - p}{k}$ is \kl{represented} by a \kl{star-free $\Rel$-polyregular function}
	defined by hardcoding the output values (in $\Rel$) when $0 \leq X \leq p$, and
	using a \kl{star-free $\Nat$-polyregular function} when $X > p$
	(\cref{binomial-function-star-free:lem}).
	Because
	$\ZSF$ is closed under products and $\Rel$-linear combinations, we conclude.
\end{proof}

\AP Obtaining an analogue of \cref{integer-binomial-polynomials:cor}
for \kl{$\Nat$-polyregular functions} requires a bit more work,
as polynomials in $\Rat[\vec{X}]$ that are \kl{represented} by $\NPoly$ are not
exactly \kl{natural binomial polynomials}
(\cref{natural-binomial-polynomial-negative:ex,natural-binomial-polynomial-basis:ex,natural-binomial-polynomial-positive-bad:ex}).
To address the issues raised by these examples, we introduce the notion of
\intro{strongly natural binomial polynomials}, as the polynomials $P \in
\Rat[X]$ such that for all partial valuation $\nu \colon \topartial \Nat$,
$\restr{P}{\nu}$ is a \kl{natural binomial polynomial}, which 
characterizes the class of polynomials that are \kl{represented} by $\NPoly$
(\cref{decide-rat-poly-npoly:cor}).

\begin{example}
	\label{natural-binomial-polynomial-negative:ex}
	The polynomial $\pbinom{X - 2}{2}$ takes negative values for $X = 0$.
\end{example}

\begin{example}
	\label{natural-binomial-polynomial-basis:ex}
	The polynomial $\pbinom{X - 2}{2} + 1$ is a \kl{non-negative} \kl{natural binomial polynomial}
	but in the standard basis, it is written
	$\pbinom{X}{2} - \pbinom{X}{1} + 2$, which has a negative coefficient.
\end{example}

\begin{example}
	\label{natural-binomial-polynomial-positive-bad:ex}
	The polynomial
	$Q(X,Y,Z) \defined \pbinom{X - 4}{1} \pbinom{Y}{1} \pbinom{Z}{1}
		+ 8 \pbinom{Y}{2} + 8 \pbinom{Z}{2} + 4$
	is a \kl{non-negative} \kl{natural binomial polynomial},
	but cannot be computed by a \kl{star-free $\Nat$-polyregular function}.
\end{example}

\begin{theorem}
	\label{decide-rat-poly-npoly:cor}
	Let $P \in \Rat[\vec{X}]$ be a polynomial with \emph{rational}
	coefficients and let $\alpha$ be the smallest number in $\Nat_{\geq 1}$
	such that $\alpha P \in \Rel[\vec{X}]$. Then, 
    the following are equivalent:
	\begin{enumerate}
		\item \label{rat-npoly-1:item} $P$ is \kl{represented} by a \kl{$\Nat$-polyregular function},
		\item \label{rat-npoly-2:item} $P$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function},
		\item \label{rat-npoly-3:item}
		      $P$ is a \kl{strongly natural binomial polynomial},
        \item \label{rat-npoly-4:item} $\alpha P \in \CorrectPoly$ 
            and $P$ is \kl{integer-valued}.
	\end{enumerate}
	In particular, the properties are decidable.
\end{theorem}
\begin{proof}

    For the implication \cref{rat-npoly-1:item} $\implies$
    \cref{rat-npoly-4:item}, we obtain $\alpha P \in \CorrectPoly$ via
    \cref{corrected-version:thm} by remarking that the \kl{star-free
    $\Nat$-polyregular functions} have output in $\Nat$ and are closed under
    multiplication by a constant $\alpha \in \Nat$. The fact that $P$ is
    \kl{integer-valued} follows from
    \cref{integer-binomial-polynomials:cor} and the fact that
    $\NSF \subseteq \ZPoly$.

    Then, \cref{rat-npoly-4:item} $\implies$
    \cref{rat-npoly-3:item}, because if $\alpha P \in \CorrectPoly$,
    it is represented by a \kl{star-free $\Nat$-polyregular function} $f$ by
    \cref{corrected-version:thm}. Let us now use \cref{n-sf-combinatorics:lem}
    to obtain a number $s \geq 1$ (that only depends on $P$) such
    that for all \kl{polynomial function} $q \colon \Nat^k \to \Sigma^*$, $g
    \circ q(X_1 + s, \dots, X_k + s)$ is a \kl{natural binomial function} $F$
    over $\Nat^k$. Let us write the
    \kl{natural binomial function} $F$ as follows $F(X_1, \dots, X_k) =
    \sum_{\card{\alpha} \leq d_2} e_{\alpha} \prod_{i = 1}^k \binom{X -
    p_{\alpha,i}}{\alpha(i)}$, with $e_{\alpha} \in \Nat$ for all
    $\card{\alpha} \leq d_2$.
    Now, because of the equality 
    $\binom{X - p - 1}{k} + \binom{X - p - 1}{k - 1} = \binom{X - p}{k}$,
    we conclude that we can find a $K \in \Nat$ and coefficients
    $f_{\alpha} \in \Nat$ such that 
    $F(X_1, \dots, X_k) =
    \sum_{\card{\alpha} \leq d_2} f_{\alpha} \prod_{i = 1}^k \binom{X_i - K}{\alpha(i)}$.
    Now, remark that $P$ is an \kl{integer-valued} polynomial of
    degree $d$, and therefore one can write $P = \sum_{\card{\alpha} \leq d}
    c_{\alpha} \prod_{i = 1}^k \pbinom{X_i - K}{\alpha(i)}$ with $c_{\alpha}
    \in \Rel$ for all $\card{\alpha} \leq d$
    (\cref{integer-binomial-polynomial:lem}). 
    As a consequence,
    $\lambda P = \sum_{\card{\alpha} \leq d} \lambda c_{\alpha} \prod_{i = 1}^k \pbinom{X_i - K - s}{\alpha(i)}$,
    and 
    by identifying coefficients in the previous two expressions, we obtain
    $c_{\alpha} = \lambda f_{\alpha}$ for all $\card{\alpha} \leq d$.
    Finally,
    $\lambda P(x_1 + s, \dots, x_k + s) = F(x_1, \dots, x_k)$ for all $\vec{x} \in \Nat^k$,
    and therefore
    $\lambda P (X_1 + s, \dots, X_k + s) =
    \sum_{\card{\alpha} \leq d_2} f_{\alpha} \prod_{i = 1}^k \pbinom{X_i - K}{\alpha(i)}$.
    In particular,
    we get
    \begin{equation*}
        \sum_{\card{\alpha} \leq d} \lambda c_{\alpha} \prod_{i = 1}^k \pbinom{X_i - K}{\alpha(i)} =
        \sum_{\card{\alpha} \leq d_2} f_{\alpha} \prod_{i = 1}^k \pbinom{X_i - K}{\alpha(i)}
    \end{equation*}
    And we conclude by term identification
    by term identification
    that $d = d_2$, and for all $\card{\alpha} \leq d$, $\lambda c_{\alpha} = f_{\alpha}$.
    As a consequence, $c_{\alpha} \in \Nat$ for all $\card{\alpha} \leq d$,
    and $P$
    is a \kl{natural binomial polynomial}.
    Remark that this reasoning remains true whenever we partially apply $P$ to some values
    in $\Nat$, as it amounts to 
    partially applying the function $f$. We have proven that 
    $P$ is a \kl{strongly natural binomial polynomial}.

    Let us now prove by induction on the number of indeterminates that
    \cref{rat-npoly-3:item} $\implies$
    \cref{rat-npoly-2:item}. Note that by construction, there exists a
    number $K \in \Nat$ such that when the input values of $P$ are all greater
    than $K$, $P$ coincides with a \kl{natural binomial function}, which is
    itself \kl{represented} by a \kl{star-free $\Nat$-polyregular function}. If
    some input value $X_i$ is set to a number $x_i \leq K$, then one can
    leverage the fact that $\restr{P}{X_i = x_i}$ remains a \kl{strongly
    natural binomial polynomial} to conclude by induction that $\restr{P}{X_i =
    x_i}$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function}.
    Combining these, we obtain a \kl{star-free $\Nat$-polyregular function}
    \kl{representing} $P$.

    Finally, the implication  \cref{rat-npoly-2:item} $\implies$ \cref{rat-npoly-1:item} is
    immediate as $\NSF \subseteq \NPoly$.
\end{proof}

Let us remark that \cref{decide-rat-poly-npoly:cor} shows that the
class of polynomials \kl{represented} by $\NPoly$ is the same as the class of
polynomials \kl{represented} by $\NSF$, which is a non-trivial statement that
will be reused in the study of more general \kl{commutative} functions in
$\ZPoly$.