### 2024-09-05 parts/conclusion.tex

%! TeX program = xelatex
%! lang = en-US
\section{Outlook}
\label{sec:ccl}

Let us conclude with a more general remark on the status of commutative input
in the study of unary output polyregular functions. Throughout the paper, we
have tried to highlight that results on polyregular functions often have the
following form: a polyregular function $f \colon \Sigma^* \to \Gamma^*$
satisfies property $X$ if and only if for all \kl{commutative} \kl{star-free
polyregular function} $g \colon \Delta^* \to \Sigma^*$, $f \circ g$ satisfies
$X$. Let us formally restate this for two earlier results in the case of
\kl{$\Rel$-polyregular functions}
in \cref{pre-compose-growth-commut:lemma,pre-compose-sf-commut:lemma}.

\begin{lemma}
    \proofref{pre-compose-growth-commut:lemma}
    \label{pre-compose-growth-commut:lemma}
    Let $f \in \ZRat$, and $k \in \Nat$. Then,
    $f \in \ZPoly[k]$ if and only if 
    for every \kl{commutative} \kl{star-free polyregular function} $h$
            of \kl{growth rate} $l \in \Nat$,
            $(f \circ h) \in \ZPoly[k\times l]$.
\end{lemma}

\begin{lemma}
    \proofref{pre-compose-sf-commut:lemma}
    \label{pre-compose-sf-commut:lemma}
    Let $f \in \ZPoly$. Then, $f \in \ZSF$,
    if and only if for every \kl{commutative} \kl{star-free polyregular function} $h$,
            $(f \circ h) \in \ZSF$.
\end{lemma}

Remark that if \cref{pre-compose-sf-commut:lemma} were to hold for
\kl{$\Nat$-polyregular functions}, then the decidability of $\NPoly$ inside
$\ZPoly$ would follow immediately, and $\NSF$ inside $\NPoly$ would also
immediately follow. On the one hand, one can guess a candidate function and
check for equivalence, on the other hand, one can guess a \kl{commutative}
\kl{star-free polyregular function} and check membership (which is decidable
thanks to this paper). This is restated in our concluding conjecture.

\begin{conjecture}
    \label{npoly-zpoly:conjecture}
    Let $k \in \Nat$.
    A function $f \in \ZPoly[k]$
    belongs to $\NPoly[k]$ if and only if
    it is \kl{$(k,\Nat)$-combinatorial}.
    In particular,
    $\NPoly[k] = \ZPoly[k] \cap \NPoly$.
\end{conjecture}