### 2024-09-16 parts/conclusion.tex
typos

%! TeX program = xelatex
%! lang = en-US
\section{Outlook}
\label{sec:ccl}

Let us end on a more general discussion regarding the status of commutative
input functions in the study of unary output polyregular functions. Throughout
the paper, we have tried to highlight that results on polyregular functions
often have the following form: a polyregular function $f \colon \Sigma^* \to
\Gamma^*$ satisfies property $X$ if and only if for all \kl{commutative}
\kl{star-free polyregular function} $g \colon \Delta^* \to \Sigma^*$, $f \circ
g$ satisfies $X$. Let us formally restate this for two earlier results in the
case of \kl{$\Rel$-polyregular functions} in
\cref{pre-compose-growth-commut:lemma,pre-compose-sf-commut:lemma}.

\begin{lemma}[restate=pre-compose-growth-commut:lemma,label=pre-compose-growth-commut:lemma]
    \proofref{pre-compose-growth-commut:lemma}
    Let $f \in \ZRat$, and $k \in \Nat$. Then,
    $f \in \ZPoly[k]$ if and only if 
    for every \kl{commutative} \kl{star-free polyregular function} $h$
            of \kl{growth rate} $l \in \Nat$,
            $(f \circ h) \in \ZPoly[k\times l]$.
\end{lemma}

\begin{lemma}[restate=pre-compose-sf-commut:lemma,label=pre-compose-sf-commut:lemma]
    \proofref{pre-compose-sf-commut:lemma}
    Let $f \in \ZPoly$. Then, $f \in \ZSF$,
    if and only if for every \kl{commutative} \kl{star-free polyregular function} $h$,
            $(f \circ h) \in \ZSF$.
\end{lemma}

Remark that if \cref{pre-compose-sf-commut:lemma} were to hold for
\kl{$\Nat$-polyregular functions}, then the decidability of $\NPoly$ inside
$\ZPoly$, and the decidability of $\NSF$ inside $\NPoly$ would immediately
follow. On the one hand, one can guess a candidate function and check for
equivalence, on the other hand, one can guess a \kl{commutative} \kl{star-free
polyregular function} and check membership (which is decidable thanks to this
paper). This is restated in our concluding conjecture.

\begin{conjecture}
    \label{npoly-zpoly:conjecture}
    Let $k \in \Nat$.
    A function $f \in \ZPoly[k]$
    belongs to $\NPoly[k]$ if and only if
    it is \kl{$(k,\Nat)$-combinatorial}.
\end{conjecture}