### 2024-11-26 pnrs.tex

% ! TeX program = xelatex
% ! TeX root = pnrs.tex
% ! lang = en-US
\documentclass[a4paper,UKenglish,autoref,thm-restate,anonymous]{lipics-v2021}
%\pdfoutput=1

% INPUTS
\input{globals/packages}
\input{globals/colorscheme}

\hideLIPIcs
%\nolinenumbers

% CONFIG
\input{globals/macros}
\input{globals/mathdefs}
\input{globals/knowledges.kl}

\crefname{conjecture}{Conjecture}{Conjectures}
\Crefname{conjecture}{Conjecture}{Conjectures}
\crefname{fact}{Fact}{Facts}
\Crefname{fact}{Fact}{Facts}
\crefname{algocf}{Algorithm}{Algorithms}
\Crefname{algocf}{Algorithm}{Algorithms}

% DOCUMENT
\begin{document}
\input{globals/metadata}

% FRONT MATTER
\maketitle
\makeabstract
%\acknowledge

%\input{parts/intro.tex}
\input{parts/newintro.tex}
\input{parts/prelim.tex}
\input{parts/counter.tex}
\input{parts/correction.tex}
\input{parts/starfree.tex}
%\input{parts/transducers.tex}
\input{parts/conclusion.tex}

% BACKMATTER
\bibliographystyle{plainurl}
\bibliography{globals/papers.bib}
%\bibliography{transducer-bib/automata,%
%transducer-bib/model-theory,%
%transducer-bib/sublinear,%
%transducer-bib/rational-series,%
%transducer-bib/polyregular,%
%transducer-bib/wqos}

\appendix

\include{parts/appendix}

\end{document}