### 2025-09-25 parts/newintro.tex
fix: pandoc + new intro

%! TeX program = xelatex
%! lang = en-US
% MAIN MATTER
\section{Introduction}\label{introduction:sec}

\AP\ Given a semiring $\mathbb{S}$, and a finite alphabet $\Sigma$, the class
of \intro{(noncommutative) $\mathbb{S}$-rational series} is defined as
functions from $\Sigma^*$ to $\mathbb{S}$ that are computed by
\kl{$\mathbb{S}$-weighted automata}~\cite{BERE10}. This computational model is
a generalization of the classical notion of non-deterministic finite automata
to the weighted setting, where transitions are labeled with elements of
$\mathbb{S}$. The semantics of \kl{$\mathbb{S}$-weighted automata} on a given
word $w$ is defined by the sum over all accepting runs reading $w$, of the
product of the weights of the transitions taken along this run. When the
semiring $\mathbb{S}$ is the Boolean semiring $\mathbb{B}$, one recovers the
classical class of regular languages via non-deterministic finite automata.

\begin{example}\label{ab-ba-power-two:ex}
  The function $f \colon \set{a,b}^* \to \Nat$ that maps a word $w$ to
  $2^{|w|_a - |w|_b} + 2^{|w|_b - |w|_a}$ is a
  $\Rat$-rational series.
  It can be computed by the following $\Rat$-weighted automaton.
  \begin{center}
    \begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
      % initial state q_0 edge from above it
      \node[state,initial above
      ] (q_0)   {$q_0$};
      \node[state,accepting] (q_1) [right=of q_0] {$q_1$};
      \node[state,accepting] (q_2) [left=of q_0] {$q_2$};
      \path[->]
        (q_0) edge [bend left] node {$a \mid 2$} (q_1)
              edge [bend left] node {$a \mid 0.5$} (q_2)
              edge [bend right] node [swap] {$b \mid 2$} (q_2)
              edge [bend right] node [swap] {$b \mid 0.5$} (q_1)
        (q_1) edge [loop above] node {$a \mid 2$} ()
              edge [loop below] node {$b \mid 0.5$} (q_0)
        (q_2) edge [loop above]  node {$b \mid 2$} ()
              edge [loop below] node {$a \mid 0.5$} ()
              ;
        \node[draw=none] (outq1) [right=1cm of q_1] {$1$};
        \node[draw=none] (outq2) [left=1cm of q_2] {$1$};
        \node[draw=none] (outq0) [below=1cm of q_0] {$1$};
        \path[->]
          (q_1) edge [] (outq1)
          (q_2) edge [] (outq2)
          (q_0) edge [] (outq0)
          ;
    \end{tikzpicture}
  \end{center}
\end{example}

\AP\ When the semiring $\mathbb{S}$ is a commutative field, for instance in the
case of the rational numbers $\Rat$, the class of \kl{$\mathbb{S}$-rational
series} is well understood, and enjoys many decidability properties by virtue
of its correspondence with \emph{linear representations}~\cite{BERE10}.
From this correspondence, one can for instance decide whether two
$\Rat$-rational series are equal, or prove that a 
$\Real$-rational series that has values in $\Rat$ is in fact a 
$\Rat$-rational series.
However, the situation is much less clear when $\mathbb{S}$ is merely a
commutative ring or worse, a commutative semiring: the corresponding theories
of \emph{modules} or \emph{semimodules} are much more involved than the one for
vector spaces, and many questions are open. For instance, the function $f$ from
\cref{ab-ba-power-two:ex} cannot be computed by a $\Nat$-weighted automaton,
even though it is a $\Rat$-rational series that only takes natural values. Even
more, we can prove that $f$ cannot be computed by any $\Rel$-weighted
automaton.

In this paper, we are interested in the case where $\mathbb{S}$ equals $\Nat$
(semiring) or $\Rel$ (ring), hence, in $\Nat$-rational series ($\intro*\NRat$)
and $\Rel$-rational series ($\intro*\ZRat$). It is clear that $\NRat$ is a
proper subclass of $\ZRat$, and a longstanding open problem is to provide an
algorithm that decides whether a given $\ZRat$ is in $\NRat$~\cite{KARH77}.

\begin{problem}\label{n-in-z-rat:problem}
    Input: A $\ZRat$ $f$. Output: Is $f$ in $\NRat$?
\end{problem}

A similar open problem has been studied in the context of rational weighted
automata, where it amounts to deciding whether a given $\Rat$-weighted
automaton is equivalent to a $\Rat_+$-weighted automaton. Let us briefly
explain why this problem is equivalent to \cref{n-in-z-rat:problem}. First,
given a $\Rat$-weighted automaton $A$, one can construct a $\Rel$-weighted
automaton $A'$ by multiplying all weights by the least common multiple of the
denominators of the weights of $A$. Semantically, $A'(w) = c^{|w|} A(w)$ for
some constant $c \in \Nat$. If $A'$ is equivalent to an $\Nat$-weighted
automaton $B$, then $A$ is $c^{-|w|} B(w)$, hence equivalent to a
$\Rat_+$-weighted automaton. For the converse implication, if $A$ is a
$\Rat$-weighted automaton equivalent to a $\Rat_+$-weighted automaton, then
$A'$ is also equivalent to some $\Rat_+$-weighted automaton $B'$, but all the
outputs of $B'$ are natural numbers, and because $\Rat_+$ is a Fatou extension
of $\Nat$,\footnote{rework this} one concludes that $B'$ is equivalent to an
$\Nat$-weighted automaton $B''$~\cite{BERE10}.

\Cref{n-in-z-rat:problem} recently received attention in the context of
\kl{polyregular functions} ($\intro*\Poly$), a computational model that aims to
generalize the theory of regular languages to the setting of string-to-string
functions~\cite{BOJA18}. In the case of regular languages, \emph{star-free
languages} form a robust subclass of regular languages described equivalently
in terms of first order logic \cite{MNPA71}, counter-free automata
\cite{MNPA71}, or aperiodic monoids \cite{SCHU65}. Analogously, there exists a
\emph{star-free} fragment of \kl{polyregular functions} called \kl{star-free
polyregular functions} ($\intro*\SF$) \cite{BOJA18}. One open question in this area is to
decide whether a given \kl{polyregular function} is \kl{star-free}.

\begin{problem}
    \label{sf-polyregular:problem}
    Input: A \kl{polyregular function} $f$. Output: Is $f$ \kl{star-free}?
\end{problem}

\AP In order to approach decision problems on \kl{polyregular functions},
restricting the output alphabet to a single letter has proven to be a fruitful
method~\cite{DOUE21,DOUE22}. Because words over a unary alphabet are
canonically identified with natural numbers, unary output \kl{polyregular
functions} are often called \kl{$\Nat$-polyregular functions} ($\NPoly$), and
their \emph{star-free} counterpart \kl{star-free $\Nat$-polyregular functions}
($\NSF$). Coincidentally, \kl{polyregular functions} with unary output forms a
subclass of \kl{$\Nat$-rational series}, namely the class of
\kl{$\Nat$-rational series} of \emph{polynomial growth}, i.e. the output of the
function is bounded by a polynomial in the size of the input 
\cite{SCHU62}.
In \cite{CDTL23},
the authors introduced the class of \kl{$\Rel$-polyregular functions}
($\ZPoly$) as a subclass of \kl{$\Rel$-rational series} that generalizes
\kl{$\Nat$-polyregular functions} by allowing negative outputs, and showed that
membership in the \emph{star-free} subclass $\ZSF$ inside $\ZPoly$ is decidable
\cite[Theorem V.8]{CDTL23}. Although this could not be immediately leveraged to
decide $\NSF$ inside $\NPoly$, it was conjectured that $\NPoly \cap \ZSF =
\NSF$ \cite[Conjecture 7.61]{DOUE23}. It was believed that understanding the
membership problem of $\NPoly$ inside $\ZPoly$, that is, a restricted version
of \cref{n-in-z-rat:problem}, would be a key step towards proving $\NPoly \cap
\ZSF = \NSF$, which itself would give hope in designing an algorithm for
\cref{sf-polyregular:problem}. We illustrate in
\cref{previously-known-inclusions:fig} the known inclusions and related open
problems between the discussed classes of functions.

\begin{figure}
    \centering
    \includestandalone[height=5cm]{tikz/class-inclusions}
    \caption{
        Decidability and inclusions of classes of functions,
        arranged along two axes. The first one is the complexity
        of the output alphabet ($\Rel$, $\Nat$, $\Sigma$). The second
        one is the allowed computational power
        (\kl{star-free polyregular functions}, \kl{polyregular functions}, 
        \kl{rational series}).
        Arrows denote strict inclusions,
        and effectiveness (both in terms of decidability and of effective
        representation) is represented by thick double arrows. Inclusions that are
        suspected to be effective are represented using a dashed arrow together with a
        question mark.
    }
    \label{previously-known-inclusions:fig}
\end{figure}

\subparagraph*{Contributions.} In this paper, we work under the extra
assumption of \emph{commutativity}, that is, assuming that the function is
invariant under the permutation of its input. In this setting, we 
prove that $\NPoly \cap \ZSF = \NSF$
\cite[Conjecture 7.61]{DOUE23} and design an algorithm that decides whether a
function in $\ZPoly$ is in $\NPoly$ \cite[Open question 5.55]{DOUE23}.
As a consequence, the
upper left square of \cref{previously-known-inclusions:fig} has all of its
arrows decidable and with effective conversion procedures under this extra
assumption. Because \kl{$\Rel$-rational series} with \emph{polynomial growth}
are exactly \kl{$\Rel$-polyregular functions} \cite{CDTL23}, this can be seen
as decision procedure for \cref{n-in-z-rat:problem} under the extra assumption
of \emph{commutativity} and \emph{polynomial growth}. Similarly, our results
provide an algorithm for \cref{sf-polyregular:problem} under the extra
assumption of \emph{commutativity} and \emph{unary output alphabet}.

As an intermediate step, we provide a complete and decidable characterization
of polynomials in $\Rat[\vec{X}]$ that can be computed using $\NRat$ (resp.
$\ZRat)$. These characterizations uncover a fatal flaw in the proof of a former
characterization of such polynomials \cite[Theorem 3.3, page 4]{KARH77}. We
also prove that this previous results holds for polynomials with at most two
indeterminates (\cref{lem:correct-covered-2}), which may explain why it was not
detected earlier. Furthermore, these characterizations provide effective
descriptions of polynomials that can be expressed in $\ZRat$ as those obtained
using integer combinations of products of \emph{binomial coefficients} (called
\kl{integer binomial polynomials}, defined page \kpageref{integer binomial
polynomial}) and similarly for $\NRat$ by introducing the notion of
\kl{strongly natural binomial polynomials} (defined page \kpageref{strongly
natural binomial polynomial}), which we believe has its own interest. In turn,
these characterizations demonstrate that polynomials expressible by $\ZRat$
(resp. $\NRat$) are exactly those expressible by $\ZSF$ (resp. $\NSF$), that
is, polynomials are inherently \emph{star free} functions.

Finally, we show that non-commutative polyregular functions in $\NPoly$ (resp.
$\NSF$) have a canonical representation in terms of their \emph{residuals}, and
propose a definition of star-free $\Rel$-rational series and $\Nat$-rational
series, that we believe is of independent interest.

\subparagraph*{Outline of the paper.} In \cref{preliminaries:sec}, we provide a
combinatorial definition of \kl{$\Nat$-polyregular functions} (resp.
\kl{$\Rel$-polyregular functions}), show that one can decide if a function $f
\in \ZPoly$ is \kl{commutative} (\cref{decidable-commutative-poly:lemma}). In
\cref{polynomials:sec}, we provide a counterexample to the flawed result of
\cite[Theorem 3.3, page 4]{KARH77} (\cref{thm:counter-example}), and correct it
by providing effective characterizations of polynomials computed by $\ZRat$
(\cref{integer-binomial-polynomials:cor}) and $\NRat$
(\cref{decide-rat-poly-npoly:cor}). Finally, in \cref{beyond-polynomials:sec},
we answer positively to \cite[Open question 5.55]{DOUE23}
(\cref{decidable-n-poly:thm}) and to \cite[Conjecture 7.61]{DOUE23}
(\cref{zsf-npoly-nsf:thm}), both under the extra assumption of
\emph{commutativity}.